---
title: "Testing Color Palettes for 2-5 groups"
author: "Dave Bosworth"
date: "2/26/2020"
output:
  html_document: 
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Create Palettes

Load R packages necessary for this document:
```{r load packages, message = FALSE, warning = FALSE}
library(colorspace)
```

Define palettes for 2-5 colors:
```{r define palettes}
pal_two <- c("#00588B", "#F69422")
pal_three <- sequential_hcl(3, palette = "Viridis")
pal_four <- sequential_hcl(4, palette = "Viridis")
pal_five <- sequential_hcl(5, palette = "Viridis")
```

# Example Plots for all Palettes

## Two Colors

Actual color Palette:
```{r two colors actual}
demoplot(pal_two, "spine")
title(main = "Two-color Palette")
```

Color vision deficiency simulations:
```{r two colors cvd simul}
col_pal <- pal_two
par(mfrow = c(2, 2))
demoplot(desaturate(col_pal), "spine")
title(main = "Desaturated")
demoplot(deutan(col_pal), "spine")
title(main = "Deuteranope")
demoplot(protan(col_pal), "spine")
title(main = "Protanope")
demoplot(tritan(col_pal), "spine")
title(main = "Tritanope")
```

## Three Colors

Actual color Palette:
```{r three colors actual}
demoplot(pal_three, "spine")
title(main = "Three-color Palette")
```

Color vision deficiency simulations:
```{r three colors cvd simul}
col_pal <- pal_three
par(mfrow = c(2, 2))
demoplot(desaturate(col_pal), "spine")
title(main = "Desaturated")
demoplot(deutan(col_pal), "spine")
title(main = "Deuteranope")
demoplot(protan(col_pal), "spine")
title(main = "Protanope")
demoplot(tritan(col_pal), "spine")
title(main = "Tritanope")
```

## Four Colors

Actual color Palette:
```{r four colors actual}
demoplot(pal_four, "spine")
title(main = "Four-color Palette")
```

Color vision deficiency simulations:
```{r four colors cvd simul}
col_pal <- pal_four
par(mfrow = c(2, 2))
demoplot(desaturate(col_pal), "spine")
title(main = "Desaturated")
demoplot(deutan(col_pal), "spine")
title(main = "Deuteranope")
demoplot(protan(col_pal), "spine")
title(main = "Protanope")
demoplot(tritan(col_pal), "spine")
title(main = "Tritanope")
```

## Five Colors

Actual color Palette:
```{r five colors actual}
demoplot(pal_five, "spine")
title(main = "Five-color Palette")
```

Color vision deficiency simulations:
```{r five colors cvd simul}
col_pal <- pal_five
par(mfrow = c(2, 2))
demoplot(desaturate(col_pal), "spine")
title(main = "Desaturated")
demoplot(deutan(col_pal), "spine")
title(main = "Deuteranope")
demoplot(protan(col_pal), "spine")
title(main = "Protanope")
demoplot(tritan(col_pal), "spine")
title(main = "Tritanope")
```


