---
title: "Testing Color Palettes"
author: "Dave Bosworth"
date: "12/24/2019"
output:
  html_document: 
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import and Prepare Data

Load R packages necessary for this analysis:
```{r load packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(scales)
library(colorspace)
```

Load common functions:
```{r load common functions}
# Load common functions
source("inlet_outlet_common_functions.R")
```

Import and Clean Load Data:
```{r import and clean load data, message = FALSE, warning = FALSE}
# Import Load Data
loads_orig <- read_csv("Loads/All_YB_Loads-R.csv")

# Clean loads_orig
loads_clean1 <- loads_orig %>% 
  # only include inlet stations and THg, MeHg
  filter(
    LocType == "Inlet",
    str_detect(Analyte, "^THg|^MeHg")
  ) %>% 
  select(
    SamplingEvent,
    StationName,
    Analyte,
    Load,
    LoadUnits
  )
```

Sum CCSB Loads:
```{r sum ccsb loads}
# Sum the CCSB Loads
ccsb <- loads_clean1 %>% 
  filter(str_detect(StationName, "^CCSB")) %>% 
  group_by(SamplingEvent, Analyte, LoadUnits) %>% 
  summarize(Load = sum(Load)) %>% 
  ungroup() %>% 
  mutate(StationName = "CCSB")

# Add back the summed CCSB loads
loads_clean2 <- loads_clean1 %>% 
  filter(!str_detect(StationName, "^CCSB")) %>%
  bind_rows(ccsb)

# Clean up
rm(ccsb)
```

Prepare Inlet Load data for Example Plots:
```{r prepare data for plots}
loads_clean <- loads_clean2 %>% 
  # rename a few stations and apply factor order
   mutate(
    StationName = case_when(
      StationName == "Knights Landing Ridge Cut" ~ "KLRC",
      StationName == "Putah Creek at Mace Blvd" ~ "Putah Ck",
      StationName == "Sac River above the Sacramento Weir" ~ "Sac Weir",
      TRUE ~ StationName
    ),
    StationName = factor(StationName, levels = c("KLRC", "CCSB", "Putah Ck", "Sac Weir", "Fremont Weir"))
   ) %>% 
  # apply factor order to SamplingEvent variable
  conv_fact_samplingevent()

# Remove intermediate df's
rm(loads_clean1, loads_clean2)
```


# Create Base Plots

Stacked Bar Plots:
```{r base plot stacked}
plot_stacked <- loads_clean %>% 
  ggplot(
    aes(
      x = SamplingEvent,
      y = Load,
      fill = StationName
    )
  ) +
  geom_col() +
  facet_wrap(
    vars(Analyte, LoadUnits),
    ncol = 3, 
    scales = "free_y",
  ) +
  labs(
    title = NULL,
    x = "Sampling Event",
    y = "Loads"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Filled Bar Plots:
```{r base plot filled}
plot_filled <- loads_clean %>% 
  ggplot(
    aes(
      x = SamplingEvent,
      y = Load,
      fill = StationName
    )
  ) +
  geom_col(position = "fill") +
  facet_wrap(
    vars(Analyte, LoadUnits),
    ncol = 3, 
    scales = "free_y",
  ) +
  labs(
    title = NULL,
    x = "Sampling Event",
    y = "Percentage of Total Inlet Load"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_continuous(labels = percent_format())
```


# Example Plots

## Viridis

Stacked Bar Plot:
```{r plot stacked viridis, fig.width = 10, fig.height = 6}
plot_stacked + scale_fill_discrete_sequential(palette = "Viridis", nmax = 5)
```

Filled Bar Plot:
```{r plot filled viridis, fig.width = 10, fig.height = 6}
plot_filled + scale_fill_discrete_sequential(palette = "Viridis", nmax = 5)
```


## Yellow-Green-Blue (YlGnBu)

Stacked Bar Plot:
```{r plot stacked YlGnBu, fig.width = 10, fig.height = 6}
plot_stacked + scale_fill_discrete_sequential(palette = "YlGnBu", nmax = 6, order = 2:6)
```

Filled Bar Plot:
```{r plot filled YlGnBu, fig.width = 10, fig.height = 6}
plot_filled + scale_fill_discrete_sequential(palette = "YlGnBu", nmax = 6, order = 2:6)
```


## Yellow-Orange-Red (YlOrRd)

Stacked Bar Plot:
```{r plot stacked YlOrRd, fig.width = 10, fig.height = 6}
plot_stacked + scale_fill_discrete_sequential(palette = "YlOrRd", nmax = 6, order = 2:6)
```

Filled Bar Plot:
```{r plot filled YlOrRd, fig.width = 10, fig.height = 6}
plot_filled + scale_fill_discrete_sequential(palette = "YlOrRd", nmax = 6, order = 2:6)
```

