---
title: "Inlet Loads to the Yolo Bypass"
author: "Dave Bosworth"
date: "2/3/2020"
output: 
  html_document: 
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import and Prepare Data

Load R packages necessary for this analysis:
```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(rlang)
library(readxl)
library(scales)
library(colorspace)
library(DT)
```

Load common functions:
```{r load common functions}
# Load common functions
source("inlet_outlet_common_functions.R")
source("Summary_Stats_1or2_Groups.R")
```

Import Load Data:
```{r import load data, message = FALSE, warning = FALSE}
# Define path on SharePoint site for data
sharepoint_path <- normalizePath(
  file.path(
    Sys.getenv("USERPROFILE"),
    "California Department of Water Resources/DWR Documents - Open Water Final Report - Documents/Technical Appendices/Inlet-Outlet/Data"
  )
)  

# Import load data
loads_orig <- read_csv(file = paste0(sharepoint_path, "/All_YB_Loads-R.csv"))
```

Clean the original loads dataframe for analysis:
```{r clean load data}
loads_inlet <- loads_orig %>% 
  # only include a subset of the data
  filter(
    str_detect(Analyte, "OC$|Hg|SS$"),
    LocType == "Inlet"
  ) %>% 
  # remove a few variables
  select(-c(Year, Conc:LocType))

# Sum the CCSB Loads
ccsb_loads <- loads_inlet %>% 
  filter(str_detect(StationName, "^CCSB")) %>% 
  group_by(SamplingEvent, Analyte, LoadUnits) %>% 
  summarize(TotalLoad = sum(Load)) %>% 
  ungroup() %>% 
  rename(Load = TotalLoad) %>% 
  mutate(StationName = "CCSB")

# Add back the summed CCSB loads and prepare data for plots
loads_clean <- loads_inlet %>% 
  filter(!str_detect(StationName, "^CCSB")) %>%
  bind_rows(ccsb_loads) %>% 
  # rename some station names and add AnalyteGroup
  mutate(
    StationName = case_when(
      StationName == "Knights Landing Ridge Cut" ~ "KLRC",
      StationName == "Putah Creek at Mace Blvd" ~ "Putah Ck",
      StationName == "Sac River above the Sacramento Weir" ~ "Sac Weir",
      TRUE ~ StationName
    ),
    AnalyteGroup = case_when(
      str_detect(Analyte, "^MeHg") ~ "MeHg",
      str_detect(Analyte, "^THg") ~ "Hg",
      str_detect(Analyte, "OC$") ~ "OrgC",
      str_detect(Analyte, "SS$") ~ "SuspSolids"
    )
  ) %>% 
  # apply plot order
  conv_fact_samplingevent() %>% 
  mutate(
    StationName = factor(
      StationName,
      levels = c(
        "KLRC", 
        "CCSB", 
        "Putah Ck", 
        "Sac Weir", 
        "Fremont Weir"
      )
    )
  )
```

# Run Calculations

Calculate total inlet loads for each analyte and sampling event:
```{r calc total inlet loads}
loads_total <- loads_clean %>% 
  group_by(SamplingEvent, Analyte, LoadUnits) %>% 
  summarize(total_load = sum(Load)) %>% 
  ungroup()
```

Calculate the inlet load percentages:
```{r calc perc inlet loads}
loads_clean_perc <- loads_clean %>% 
  left_join(loads_total, by = c("SamplingEvent", "Analyte", "LoadUnits")) %>% 
  mutate(percent_load = Load/total_load) %>% 
  select(-c(Load:total_load)) %>% 
  arrange(SamplingEvent, StationName)

loads_clean_perc_wide <- loads_clean_perc %>% 
  pivot_wider(
    names_from = StationName,
    values_from = percent_load
  ) %>% 
  replace_na(list("Fremont Weir" = 0, "Sac Weir" = 0))
```

Calculate summary statistics of loads for each inlet:
```{r calc summ stats inlet loads}
loads_clean_summ <- loads_clean %>% 
  SummStat(Load, StationName, Analyte) %>% 
  mutate_at(vars(Mean:IQR), signif, digits = 3)
```

Calculate summary statistics of inlet load percentages:
```{r calc summ stats perc inlet loads}
loads_clean_perc_summ <- loads_clean_perc %>% 
  SummStat(percent_load, StationName, Analyte)
```

# Create Plot Functions

Function for facet barplots for inlet loads:
```{r fun facet barplots inlets}
barplot_inlet_facet <- function(df, ncols) {
  p <- 
    ggplot(
      data = df,
      aes(
        x = SamplingEvent,
        y = Load
      )
    ) +
    geom_col() +
    facet_wrap(
      vars(StationName, Analyte),
      scales = "free_y",
      ncol = ncols
    ) +
    labs(
      title = NULL,
      x = NULL,
      y = paste0("Loads (", df$LoadUnits[1], ")")
    ) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
  return(p)
}
```

Function for barplots showing percentages of inlet loads:
```{r fun barplots inlet perc}
barplot_inlet_percent <- function(df) {
  p <- 
    ggplot(
      data = df,
      aes(
        x = SamplingEvent,
        y = Load,
        fill = StationName
      )
    ) +
    geom_col(position = "fill") +
    facet_wrap(vars(Analyte)) +
    labs(
      title = NULL,
      x = NULL,
      y = "Percent of Inlet Load"
    ) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    scale_y_continuous(labels = percent_format()) +
    scale_fill_discrete_sequential(palette = "Viridis", nmax = 5)
  
  return(p)
}
```

Function for boxplots of inlet loads:
```{r fun boxplots inlet}
boxplot_inlet <- function(df) {
  p <- 
    ggplot(
      data = df,
      aes(
        x = StationName,
        y = Load
      )
    ) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(width = 0.25) +
    stat_summary( 
      fun.y = "mean", #add a symbol representing the mean of each group to the plot
      color = "red",
      geom = "point",
      shape = 9, #diamond cross shape
      size = 2 
    ) +
    facet_wrap(vars(Analyte)) +
    labs(
      title = NULL,
      caption = "Red diamond represents the average",
      x = NULL,
      y = paste0("Loads (", df$LoadUnits[1], ")")
    ) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
  return(p)
}
```

# Create Plots

```{r create plots}
loads_plots <- loads_clean %>% 
  group_nest(AnalyteGroup) %>% 
  mutate(
    ncols = if_else(AnalyteGroup == "SuspSolids", 2, 3),
    barplots_facet = map2(data, ncols, .f = barplot_inlet_facet),
    barplots_percent = map(data, .f = barplot_inlet_percent),
    boxplots = map(data, .f = boxplot_inlet)
  )
```

# Plots and Summary Statistics

## MeHg

### Facet Barplots

```{r mehg inlet facet barplots, fig.height = 9}
loads_plots_mehg <- filter(loads_plots, AnalyteGroup == "MeHg")

loads_plots_mehg %>% pull(barplots_facet) %>% chuck(1)
```

### Proportional Barplots

```{r mehg inlet perc barplots, fig.height = 6, fig.width = 8}
loads_plots_mehg %>% pull(barplots_percent) %>% chuck(1)
```

### Boxplots

```{r mehg inlet boxplots, fig.height = 6, fig.width = 8}
loads_plots_mehg %>% pull(boxplots) %>% chuck(1)
```

### Total Load Barplots

```{r mehg total load barplots}
loads_total_mehg <- filter(loads_total, Analyte %in% c("MeHg- filtered", "MeHg- particulate"))

loads_total_mehg %>%  
  ggplot(aes(x = SamplingEvent, y = total_load, fill = Analyte)) +
  geom_col() +
  labs(
    title = NULL,
    x = NULL,
    y = "MeHg Loads (g/day)"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

loads_total_mehg %>%  
  ggplot(aes(x = SamplingEvent, y = total_load, fill = Analyte)) +
  geom_col(position = "fill") +
  labs(
    title = NULL,
    x = NULL,
    y = "Percent of Inlet Load"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_continuous(labels = percent_format())
```

### Summary Statistics of each Inlet

```{r mehg summ stats inlets}
loads_clean_summ %>% 
  filter(str_detect(Analyte, "^MeHg")) %>% 
  datatable(rownames = FALSE)
```

### % of total inlet load for each inlet

```{r mehg perc load inlets}
loads_clean_perc_wide %>% 
  filter(str_detect(Analyte, "^MeHg")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(3:7, digits = 1)
```

### Summary Statistics of % loads

```{r mehg summ stats perc inlets}
loads_clean_perc_summ %>% 
  filter(str_detect(Analyte, "^MeHg")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(4:11, digits = 1)
```

## Hg

### Facet Barplots

```{r hg inlet facet barplots, fig.height = 9}
loads_plots_hg <- filter(loads_plots, AnalyteGroup == "Hg")

loads_plots_hg %>% pull(barplots_facet) %>% chuck(1)
```

### Proportional Barplots

```{r hg inlet perc barplots, fig.height = 6, fig.width = 8}
loads_plots_hg %>% pull(barplots_percent) %>% chuck(1)
```

### Boxplots

```{r hg inlet boxplots, fig.height = 6, fig.width = 8}
loads_plots_hg %>% pull(boxplots) %>% chuck(1)
```

### Total Load Barplots

```{r hg total load barplots}
loads_total_hg <- filter(loads_total, Analyte %in% c("THg- filtered", "THg- particulate"))

loads_total_hg %>%  
  ggplot(aes(x = SamplingEvent, y = total_load, fill = Analyte)) +
  geom_col() +
  labs(
    title = NULL,
    x = NULL,
    y = "Hg Loads (g/day)"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

loads_total_hg %>%  
  ggplot(aes(x = SamplingEvent, y = total_load, fill = Analyte)) +
  geom_col(position = "fill") +
  labs(
    title = NULL,
    x = NULL,
    y = "Percent of Inlet Load"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_continuous(labels = percent_format())
```

### Summary Statistics of each Inlet

```{r hg summ stats inlets}
loads_clean_summ %>% 
  filter(str_detect(Analyte, "^THg")) %>% 
  datatable(rownames = FALSE)
```

### % of total inlet load for each inlet

```{r hg perc load inlets}
loads_clean_perc_wide %>% 
  filter(str_detect(Analyte, "^THg")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(3:7, digits = 1)
```

### Summary Statistics of % loads

```{r hg summ stats perc inlets}
loads_clean_perc_summ %>% 
  filter(str_detect(Analyte, "^THg")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(4:11, digits = 1)
```

## Organic Carbon

### Facet Barplots

```{r oc inlet facet barplots, fig.height = 9}
loads_plots_oc <- filter(loads_plots, AnalyteGroup == "OrgC")

loads_plots_oc %>% pull(barplots_facet) %>% chuck(1)
```

### Proportional Barplots

```{r oc inlet perc barplots, fig.height = 6, fig.width = 8}
loads_plots_oc %>% pull(barplots_percent) %>% chuck(1)
```

### Boxplots

```{r oc inlet boxplots, fig.height = 6, fig.width = 8}
loads_plots_oc %>% pull(boxplots) %>% chuck(1)
```

### Total Load Barplots

```{r oc total load barplots}
loads_total_oc <- filter(loads_total, Analyte %in% c("DOC", "POC"))

loads_total_oc %>%  
  ggplot(aes(x = SamplingEvent, y = total_load, fill = Analyte)) +
  geom_col() +
  labs(
    title = NULL,
    x = NULL,
    y = "Organic Carbon Loads (1,000 kg/day)"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

loads_total_oc %>%  
  ggplot(aes(x = SamplingEvent, y = total_load, fill = Analyte)) +
  geom_col(position = "fill") +
  labs(
    title = NULL,
    x = NULL,
    y = "Percent of Inlet Load"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_continuous(labels = percent_format())
```

### Summary Statistics of each Inlet

```{r oc summ stats inlets}
loads_clean_summ %>% 
  filter(str_detect(Analyte, "OC$")) %>% 
  datatable(rownames = FALSE)
```

### % of total inlet load for each inlet

```{r oc perc load inlets}
loads_clean_perc_wide %>% 
  filter(str_detect(Analyte, "OC$")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(3:7, digits = 1)
```

### Summary Statistics of % loads

```{r oc summ stats perc inlets}
loads_clean_perc_summ %>% 
  filter(str_detect(Analyte, "OC$")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(4:11, digits = 1)
```

## Suspended Solids

### Facet Barplots

```{r ss inlet facet barplots, fig.height = 9}
loads_plots_ss <- filter(loads_plots, AnalyteGroup == "SuspSolids")

loads_plots_ss %>% pull(barplots_facet) %>% chuck(1)
```

### Proportional Barplots

```{r ss inlet perc barplots}
loads_plots_ss %>% pull(barplots_percent) %>% chuck(1)
```

### Boxplots

```{r ss inlet boxplots, fig.height = 6, fig.width = 8}
loads_plots_ss %>% pull(boxplots) %>% chuck(1)
```

### Total Load Barplots

```{r tss total load barplots}
loads_total %>% 
  filter(Analyte == "TSS") %>% 
  ggplot(aes(x = SamplingEvent, y = total_load)) +
  geom_col() +
  labs(
    title = "TSS",
    x = NULL,
    y = "TSS Loads (1,000 kg/day)"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r vss total load barplots}
loads_total %>% 
  filter(Analyte == "VSS") %>% 
  ggplot(aes(x = SamplingEvent, y = total_load)) +
  geom_col() +
  labs(
    title = "VSS",
    x = NULL,
    y = "VSS Loads (1,000 kg/day)"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Summary Statistics of each Inlet

```{r ss summ stats inlets}
loads_clean_summ %>% 
  filter(str_detect(Analyte, "SS$")) %>% 
  datatable(rownames = FALSE)
```

### % of total inlet load for each inlet

```{r ss perc load inlets}
loads_clean_perc_wide %>% 
  filter(str_detect(Analyte, "SS$")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(3:7, digits = 1)
```

### Summary Statistics of % loads

```{r ss summ stats perc inlets}
loads_clean_perc_summ %>% 
  filter(str_detect(Analyte, "SS$")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(4:11, digits = 1)
```

