---
title: 'Flood Comparison: 2006 vs 2017'
author: "Dave Bosworth"
date: "5/18/2020"
output: 
  html_document: 
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import and Prepare Data

Load R packages necessary for this analysis:
```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(readxl)
library(lubridate)
library(scales)
library(openwaterhg)
```

Bring in and prepare MeHg load and flow data for 2017 flood:
```{r clean 2017 load and flow data, message = FALSE}
# Bring in data
source("YB_Mass_Balance/Loads/Import_Net_Load_Data.R")
source("YB_Mass_Balance/Loads/Import_Total_Load_Data.R")
source("YB_Mass_Balance/Flows/Import_Inlet_Flow_Data_SE.R")

# Calculate total inflows for sampling events in 2017
flows_inlet_total17 <- flows_inlet_se %>% 
  filter(Year == 2017) %>% 
  group_by(SamplingEvent) %>% 
  summarize(total_inflow = sum(Flow))

# Clean loads_total df
loads_total_mod <- loads_total %>%
  filter(LocType != "Below Liberty") %>% 
  pivot_wider(
    id_cols = -digits,
    names_from = LocType,
    values_from = total_load
  )

# Join net and total MeHg loads and flow data
loads_umehg_flow17 <- loads_net %>% 
  filter(Reach == "Upper") %>%
  select(-c(Reach, digits)) %>% 
  left_join(loads_total_mod) %>% 
  filter(
    Year == 2017,
    Analyte == "MeHg- total"
  ) %>%
  left_join(flows_inlet_total17) %>% 
  select(Year, net_load:total_inflow) %>% 
  rename(
    Net_Load = net_load,
    Inlet_Load = Inlet,
    Outlet_Load = Outlet,
    Total_Inflow = total_inflow
  )

# Clean up
rm(flows_inlet_se, flows_inlet_total17, loads_total, loads_total_mod, loads_net)
```

Bring in and prepare MeHg load and flow data for 2006 flood:
```{r clean 2006 load and flow data}
loads_flow_cf_clean <- loads_flow_cf %>% 
  select(Year, Inlet_Load:Total_Inflow)
```

Bind MeHg load and flow data together:
```{r bind load and flow data}
loads_flow_comb <-   
  bind_rows(loads_flow_cf_clean, loads_umehg_flow17) %>% 
  pivot_longer(
    cols = Inlet_Load:Net_Load,
    names_to = "Load_Type",
    values_to = "Load"
  ) %>% 
  mutate(
    Year = as.character(Year),
    Load_Type = factor(Load_Type, levels = c("Inlet_Load", "Outlet_Load", "Net_Load"))
  )

# Clean up
rm(loads_flow_cf_clean, loads_umehg_flow17)
```

Bring in and prepare Lisbon stage data for 2017 flood:
```{r clean 2017 lis stage data, message = FALSE}
# Define path on SharePoint site for data
sharepoint_path <- normalizePath(
  file.path(
    Sys.getenv("USERPROFILE"),
    "California Department of Water Resources/DWR Documents - Open Water Final Report - Documents/Technical Appendices/Technical Appendix-B_Inlet-Outlet/Data/Raw"
  )
)

# Import 2017 Lisbon Stage Data 
lisb17_orig <- read_excel(
  paste0(sharepoint_path, "/2017_YB_Flood_Flows.xlsx"), 
  range = "Lisbon!A1:C15195"
)

# Clean 2017 Lisbon Stage Data
lisb17_clean <- lisb17_orig %>% 
  select(-Date) %>% 
  rename(
    datetime = "Date/Time (PST)",
    stage = "Stage (ft)"
  )

# Create a df of sampling events to mark these on the Lisbon hydrograph
se_datetimes17 <- tibble(
  datetime = as_datetime(
    c(
      "2017-01-11 12:00:00",
      "2017-01-24 12:00:00",
      "2017-01-31 12:00:00",
      "2017-02-14 12:00:00",
      "2017-03-01 12:00:00",
      "2017-03-15 12:00:00",
      "2017-03-28 12:00:00",
      "2017-04-11 12:00:00",
      "2017-04-25 12:00:00"
    )
  )
)

# Join se_datetimes df to lisb17_clean to pull out data for just the sampling events
se_lisb17 <- inner_join(lisb17_clean, se_datetimes17)

# Clean up
rm(se_datetimes17)
```

Bring in and prepare Lisbon stage data for 2006 flood:
```{r clean 2006 lis stage data, message = FALSE}
# Import 2006 Lisbon Stage Data 
lisb06_orig <- read_excel(
  paste0(sharepoint_path, "/2006_Lisbon_Stage_Data.xlsx"), 
  range = "A1:C17377"
)

# Clean 2006 Lisbon Stage Data
lisb06_clean <- lisb06_orig %>% 
  select(-Date) %>% 
  rename(
    datetime = "Date/Time (PST)",
    stage = "Stage (ft)"
  )

# Create a df of sampling events to mark these on the Lisbon hydrograph
se_datetimes06 <- loads_flow_cf %>% 
  select(SampleDate) %>% 
  mutate(
    SampleDate = as.character(SampleDate),
    SampleDate = paste0(SampleDate, " 12:00:00"),
    datetime = as_datetime(SampleDate)
  ) %>% 
  select(-SampleDate)

# Join se_datetimes df to lisb06_clean to pull out data for just the sampling events
se_lisb06 <- inner_join(lisb06_clean, se_datetimes06)

# Clean up
rm(se_datetimes06)
```


# Plots

## MeHg loads vs total inflow comparison

```{r mehg loads vs inflow plots, fig.width = 8.5, fig.height = 4}
loads_flow_comb %>% 
  ggplot(aes(x = Total_Inflow, y = Load, color = Year)) +
  geom_point() +
  geom_smooth(
    method = "lm",
    formula = y ~ x,
    se = FALSE
  ) +
  facet_wrap(
    vars(Load_Type),
    labeller = labeller(
      Load_Type = c(
        Inlet_Load = "Inlet Load",
        Outlet_Load = "Outlet Load",
        Net_Load = "Net Load"
      )
    )
  ) +
  xlab("Total Bypass Inflow (cfs)") +
  ylab("Load (g/day)") +
  theme_owhg() +
  add_gen_color_pal(
    num_colors = 2,
    aes_type = "color",
    legend_title = "Flood"
  ) +
  scale_x_continuous(labels = label_comma())
```


## Lisbon Hydrographs

### 2006

```{r lisb hydrograph 2006}
lisb06_clean %>% 
  ggplot(aes(x = datetime, y = stage)) +
  geom_line() +
  geom_point(
    data = se_lisb06, 
    color = "red",
    size = 2
  ) +
  ylab("Stage (ft)") +
  xlab(NULL) +
  scale_x_datetime(
    breaks = breaks_pretty(15),
    labels = label_date_short(),
    expand = c(0.01, 0.01)
  ) +
  theme_owhg()
```


### 2017

```{r lisb hydrograph 2017}
lisb17_clean %>% 
  ggplot(aes(x = datetime, y = stage)) +
  geom_line() +
  geom_point(
    data = se_lisb17, 
    color = "red",
    size = 2
  ) +
  ylab("Stage (ft)") +
  xlab(NULL) +
  scale_x_datetime(
    breaks = breaks_pretty(15),
    labels = label_date_short(),
    expand = c(0.01, 0.01)
  ) +
  theme_owhg()
```

