---
title: "Inlet Loads to the Yolo Bypass"
author: "Dave Bosworth"
date: "2/3/2020"
output: 
  html_document: 
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import and Prepare Data

Load R packages necessary for this analysis:
```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(scales)
library(DT)
library(openwaterhg)
```

Clean the original loads_calc dataframe for analysis:
```{r clean load data, message = FALSE}
# Bring in load data for the inlets
source("YB_Mass_Balance/Loads/Import_Inlet_Load_Data.R")

# Prepare inlet load data for plots and summaries
loads_inlet_clean <- loads_inlet %>% 
  # only include a subset of the data
  filter(str_detect(Analyte, "OC$|Hg|SS$")) %>% 
  # Add AnalyteGroup variable
  mutate(
    AnalyteGroup = case_when(
      str_detect(Analyte, "^MeHg") ~ "MeHg",
      str_detect(Analyte, "^THg") ~ "Hg",
      str_detect(Analyte, "OC$") ~ "OrgC",
      str_detect(Analyte, "SS$") ~ "SuspSolids"
    )
  ) %>% 
  # apply plot order
  conv_fact_samplingevent() %>% 
  conv_fact_inlet_names() %>% 
  # remove a few variables
  select(-c(Year, LocType))

# Apply factors to zero_loads df
zero_loads <- zero_loads %>% 
  conv_fact_samplingevent() %>% 
  conv_fact_inlet_names()

# Clean up
rm(loads_inlet)
```

Create a new loads_clean df with inlet loads only for 2017 sampling events. This will be used for the summary statistics and boxplots.
```{r filter 2017 data}
loads_inlet_clean17 <- loads_inlet_clean %>% filter(str_detect(SamplingEvent, "17$"))
```

Create rounded versions of the inlet load dataframes for some of the analyses:
```{r round inlet load data}
loads_inlet_clean_r <- loads_inlet_clean %>% 
  mutate(Load = if_else(!is.na(digits), signif(Load, digits), Load)) %>% 
  select(-digits)

loads_inlet_clean17_r <- loads_inlet_clean17 %>% 
  mutate(Load = if_else(!is.na(digits), signif(Load, digits), Load)) %>% 
  select(-digits)
```


Bring in total inlet loads data for each analyte and sampling event:
```{r import total inlet loads}
# Bring in total load data
source("YB_Mass_Balance/Loads/Import_Total_Load_Data.R")

# Prepare total inlet load data for plots
loads_total_clean <- loads_total %>% 
  # only include a subset of the data
  filter(
    LocType == "Inlet",
    str_detect(Analyte, "OC$|Hg|SS$")
  ) %>% 
  # Add AnalyteGroup variable
  mutate(
    AnalyteGroup = case_when(
      str_detect(Analyte, "^MeHg") ~ "MeHg",
      str_detect(Analyte, "^THg") ~ "Hg",
      str_detect(Analyte, "OC$") ~ "OrgC",
      str_detect(Analyte, "SS$") ~ "SuspSolids"
    )
  ) %>% 
  # apply plot order
  conv_fact_samplingevent() %>% 
  # remove a few variables
  select(-c(Year, LocType))

# Create a rounded version of loads_total_clean
loads_total_clean_r <- loads_total_clean %>% 
  # Round total loads to appropriate number of significant figures
  mutate(total_load = signif(total_load, digits)) %>% 
  select(-digits)

# Clean up
rm(loads_total)
```


# Run Calculations

Calculate the inlet load percentages:
```{r calc perc inlet loads, message = FALSE}
# Calculate total inlet loads of each parameter for each sampling event
loads_inlet_total <- loads_inlet_clean %>% 
  group_by(SamplingEvent, Analyte) %>% 
  summarize(total_load = sum(Load)) %>% 
  ungroup()

# Calculate percentage of total load for each tributary
loads_inlet_perc <- loads_inlet_clean %>% 
  select(-digits) %>% 
  left_join(loads_inlet_total) %>% 
  mutate(percent_load = Load/total_load) %>% 
  select(-c(LoadUnits, Load, AnalyteGroup, total_load)) %>% 
  arrange(SamplingEvent, StationName)

loads_inlet_perc_wide <- loads_inlet_perc %>% 
  pivot_wider(names_from = StationName, values_from = percent_load)

# Clean up
rm(loads_inlet_total)
```

Calculate summary statistics of loads for each inlet:
```{r calc summ stats inlet loads, message = FALSE}
# Including zero load values for Fremont and Sacramento Weirs
# Calculate averages and standard deviations on non-rounded data
loads_inlet_summ1 <- loads_inlet_clean17 %>% 
  group_by(StationName, Analyte) %>% 
  summarize(
    sign_digits = min(digits, na.rm = TRUE),
    Mean = signif(mean(Load), sign_digits),
    StDev = signif(sd(Load), sign_digits)
  ) %>% 
  ungroup() %>% 
  select(-sign_digits)

# Calculate all other summary statistics on rounded data
loads_inlet_summ2 <- loads_inlet_clean17_r %>% 
  summ_stat(Load, StationName, Analyte) %>% 
  select(-c(Mean, StDev))

# Join Summary Statistics together
loads_inlet_summ <- 
  left_join(loads_inlet_summ2, loads_inlet_summ1) %>% 
  select(StationName:N, Mean, StDev, Minimum:IQR)

# NOT Including zero load values for Fremont and Sacramento Weirs
# Calculate averages and standard deviations on non-rounded data
loads_inlet_summ_no0_1 <- loads_inlet_clean17 %>%
  anti_join(zero_loads) %>% 
  group_by(StationName, Analyte) %>% 
  summarize(
    sign_digits = min(digits, na.rm = TRUE),
    Mean = signif(mean(Load), sign_digits),
    StDev = signif(sd(Load), sign_digits)
  ) %>% 
  ungroup() %>% 
  select(-sign_digits)

# Calculate all other summary statistics on rounded data
loads_inlet_summ_no0_2 <- loads_inlet_clean17_r %>% 
  anti_join(zero_loads) %>% 
  summ_stat(Load, StationName, Analyte) %>% 
  select(-c(Mean, StDev))

# Join Summary Statistics together
loads_inlet_summ_no0 <- 
  left_join(loads_inlet_summ_no0_2, loads_inlet_summ_no0_1) %>% 
  select(StationName:N, Mean, StDev, Minimum:IQR)

# Clean up
rm(loads_inlet_summ1, loads_inlet_summ2, loads_inlet_summ_no0_1, loads_inlet_summ_no0_2)
```

Calculate summary statistics of inlet load percentages:
```{r calc summ stats perc inlet loads, message = FALSE}
# Including zero percentage values for Fremont and Sacramento Weirs
loads_inlet_perc_summ <- loads_inlet_perc %>% 
  filter(str_detect(SamplingEvent, "17$")) %>% 
  summ_stat(percent_load, StationName, Analyte) %>% 
  mutate_at(vars(Mean:IQR), signif, digits = 3)

# NOT Including zero percentage values for Fremont and Sacramento Weirs
loads_inlet_perc_summ_no0 <- loads_inlet_perc %>% 
  filter(str_detect(SamplingEvent, "17$")) %>% 
  anti_join(zero_loads) %>% 
  summ_stat(percent_load, StationName, Analyte) %>% 
  mutate_at(vars(Mean:IQR), signif, digits = 3)
```

Calculate %MeHg of Hg loads for all fractions:
```{r calc perc mehg inlet loads, message = FALSE}
loads_inlet_perc_mehg <- loads_inlet_clean %>% 
  anti_join(zero_loads) %>% 
  filter(str_detect(AnalyteGroup, "Hg$")) %>% 
  separate(Analyte, into = c("Analyte", "Fraction")) %>% 
  select(-c(LoadUnits, AnalyteGroup)) %>% 
  pivot_wider(
    id_cols = -digits,
    names_from = Analyte,
    values_from = Load
  ) %>% 
  mutate(
    per_mehg = MeHg/THg,
    Fraction = str_to_title(Fraction)
  ) %>% 
  select(-c(MeHg, THg))
```

Calculate the % of each fraction for Hg and MeHg loads:
```{r calc perc fraction hg and mehg, message = FALSE}
loads_inlet_perc_frac <- loads_inlet_clean %>% 
  anti_join(zero_loads) %>% 
  filter(str_detect(AnalyteGroup, "Hg$")) %>% 
  separate(Analyte, into = c("Analyte", "Fraction"), sep = "- ") %>% 
  pivot_wider(
    id_cols = c(SamplingEvent, StationName, Analyte),
    names_from = Fraction,
    values_from = Load
  ) %>% 
  mutate(
    perc_filt = filtered/(filtered + particulate),
    perc_part = particulate/(filtered + particulate)
  ) %>% 
  select(-c(filtered, particulate, total))
```

Calculate summary statistics of the % fractions for Hg and MeHg loads:
```{r calc summ stats perc fraction hg and mehg}
loads_inlet_perc_frac_summ <- loads_inlet_perc_frac %>% 
  # Just for the sampling events in 2017
  filter(str_detect(SamplingEvent, "17$")) %>% 
  rename(
    Filtered = perc_filt,
    Particulate = perc_part
  ) %>% 
  pivot_longer(
    cols = Filtered:Particulate,
    names_to = "Fraction",
    values_to = "perc"
  ) %>% 
  group_nest(Analyte) %>% 
  mutate(
    summ_stats = map(
      data,
      .f = ~summ_stat(.x, perc, StationName, Fraction) %>% 
        mutate_at(vars(Mean:IQR), signif, digits = 3)
    )
  )
```


# Create Plot Functions

Function for facet barplots of inlet loads, free y-axis scale:
```{r fun facet barplots inlets free}
barplot_inlet_facet_free <- function(df, ncols) {
  p <- 
    ggplot(
      data = df,
      aes(
        x = SamplingEvent,
        y = Load
      )
    ) +
    geom_col() +
    facet_wrap(
      vars(StationName, Analyte),
      scales = "free_y",
      ncol = ncols
    ) +
    labs(
      title = NULL,
      x = NULL,
      y = paste0("Loads (", df$LoadUnits[1], ")")
    ) +
    theme_owhg(x_axis_v = TRUE)
  
  return(p)
}
```

Function for facet barplots of inlet loads, fixed y-axis scale:
```{r fun facet barplots inlets fixed}
barplot_inlet_facet_fixed <- function(df, param) {
  p <- 
    ggplot(
      data = df,
      aes(
        x = SamplingEvent,
        y = Load
      )
    ) +
    geom_col() +
    facet_grid(cols = vars(StationName)) +
    labs(
      title = paste0(param, " (", df$LoadUnits[1], ")"),
      x = NULL,
      y = NULL
    ) +
    theme_owhg(x_axis_v = TRUE)
  
  return(p)
}
```

Function for barplots showing percentages of inlet loads:
```{r fun barplots inlet perc}
barplot_inlet_percent <- function(df) {
  p <- 
    ggplot(
      data = df,
      aes(
        x = SamplingEvent,
        y = Load,
        fill = StationName
      )
    ) +
    geom_col(position = "fill") +
    facet_wrap(vars(Analyte)) +
    labs(
      title = NULL,
      x = NULL,
      y = "Percent of Inlet Load"
    ) +
    scale_y_continuous(labels = percent_format()) +
    add_inlet_color_pal("fill", "Inlet") +
    theme_owhg(x_axis_v = TRUE)
  
  return(p)
}
```

Function for boxplots of inlet loads:
```{r fun boxplots inlet}
boxplot_inlet <- function(df) {
  p <- 
    ggplot(
      data = df,
      aes(
        x = StationName,
        y = Load
      )
    ) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(width = 0.25) +
    stat_summary( 
      fun = "mean", #add a symbol representing the mean of each group to the plot
      color = "red",
      geom = "point",
      shape = 9, #diamond cross shape
      size = 2 
    ) +
    facet_wrap(vars(Analyte)) +
    labs(
      title = paste0("Loads (", df$LoadUnits[1], ")"),
      x = NULL,
      y = NULL
    ) +
    theme_owhg(x_axis_v = TRUE)
    
  return(p)
}
```

Function for barplots of total inlet loads by fraction:
```{r fun barplots total inlet}
barplot_total_frac <- function(df, param) {
  p <-  
    ggplot(
      data = df,
      aes(
        x = SamplingEvent, 
        y = total_load, 
        fill = Analyte
      )
    ) +
    geom_col() +
    labs(
      title = NULL,
      x = NULL,
      y = paste0(param, " Loads (", df$LoadUnits[1], ")")
    ) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    add_gen_color_pal(2, "fill", "Analyte")
  
  return(p)
}
```

Function for proportional barplots of total inlet loads by fraction:
```{r fun barplots total inlet prop}
barplot_total_frac_prop <- function(df) {
  p <-  
    ggplot(
      data = df,
      aes(
        x = SamplingEvent, 
        y = total_load, 
        fill = Analyte
      )
    ) +
    geom_col(position = "fill") +
    labs(
      title = NULL,
      x = NULL,
      y = "Percent of Inlet Load"
    ) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    add_gen_color_pal(2, "fill", "Analyte") +
    scale_y_continuous(labels = percent_format())
  
  return(p)
}
```

Function for barplots showing %MeHg of Hg loads:
```{r fun barplots perc mehg}
barplot_perc_mehg <- function(df) {
  p <- 
    ggplot(
      data = df,
      aes(
        x = SamplingEvent,
        y = per_mehg
      )
    ) +
    geom_col() +
    facet_grid(rows = vars(StationName)) +
    labs(
      title = "Percent MeHg of Hg",
      x = NULL,
      y = NULL
    ) +
    theme_owhg(x_axis_v = TRUE)
  
  return(p)
}
```

Function to have plotly not add markers for outliers:
```{r fun remove plotly markers}
remove_markers <- function(x) {
  if (x$type == "box") {
      x$marker = list(opacity = 0)
    }
    return(x)
}
```

# Create Plots

```{r create plots, message = FALSE}
# Barplots of all individual sampling events
  # Use rounded values for the barplots
  loads_barplots_free <- loads_inlet_clean_r %>% 
    anti_join(zero_loads) %>% 
    group_nest(AnalyteGroup) %>% 
    mutate(
      ncols = if_else(AnalyteGroup == "SuspSolids", 2, 3),
      barplots_facet_free = map2(data, ncols, .f = barplot_inlet_facet_free),
    ) %>% 
    select(-data)
  
  # Use unrounded values for the proportional barplots
  loads_barplots_prop <- loads_inlet_clean %>% 
    anti_join(zero_loads) %>% 
    group_nest(AnalyteGroup) %>% 
    mutate(barplots_percent = map(data, .f = barplot_inlet_percent)) %>% 
    select(-data)

# Boxplots- just 2017 sampling events, rounded values 
loads_boxplots <- loads_inlet_clean17_r %>% 
  group_nest(AnalyteGroup) %>% 
  mutate(boxplots = map(data, .f = boxplot_inlet)) %>% 
  select(-data)

# Total load barplots
  # Use rounded values for the barplots
  loads_barplots_total <- loads_total_clean_r %>% 
    filter(str_detect(Analyte, "filtered$|particulate$|DOC|POC")) %>% 
    group_nest(AnalyteGroup) %>% 
    mutate(barplots_total = map2(data, AnalyteGroup, .f = barplot_total_frac)) %>% 
    select(-data)

  # Use unrounded values for the proportional barplots
  loads_barplots_total_prop <- loads_total_clean %>% 
    filter(str_detect(Analyte, "filtered$|particulate$|DOC|POC")) %>% 
    group_nest(AnalyteGroup) %>% 
    mutate(barplots_total_prop = map(data, .f = barplot_total_frac_prop)) %>% 
    select(-data)

# Join five dataframes together
loads_plots <- 
  reduce(
    list(
      loads_barplots_free, 
      loads_barplots_prop, 
      loads_boxplots,
      loads_barplots_total,
      loads_barplots_total_prop
    ),
    left_join
  )

# Barplots for just uHg, uMeHg, TOC, and TSS and 2017 sampling events, fixed y-axis scale
loads_barplots_fixed <- loads_inlet_clean17_r %>% 
  anti_join(zero_loads) %>% 
  group_nest(Analyte) %>% 
  filter(str_detect(Analyte, "total$|TOC|TSS")) %>% 
  mutate(
    barplots_facet_fixed = map2(data, Analyte, .f = barplot_inlet_facet_fixed)
  )

# Clean up 
rm(
  loads_barplots_free,
  loads_barplots_prop,
  loads_boxplots,
  loads_barplots_total,
  loads_barplots_total_prop
)
```


# Plots and Summary Statistics {.tabset .tabset-pills}

The summary statistics for the inlet and % loads and the boxplots include only the sampling events conducted in 2017. The remaining plots include all sampling events conducted throughout the study period (2014-2017). The boxplots include zero load values for the Fremont and Sacramento Weirs when they weren't spilling.

## MeHg {.tabset .tabset-pills}

### Facet Barplots

```{r mehg inlet facet barplots free, fig.height = 9}
loads_plots_mehg <- filter(loads_plots, AnalyteGroup == "MeHg")

loads_plots_mehg %>% pull(barplots_facet_free) %>% chuck(1)
```

```{r mehg inlet facet barplots fixed, fig.height = 5.5, fig.width = 8}
loads_barplots_fixed %>% 
  filter(Analyte == "MeHg- total") %>% 
  pull(barplots_facet_fixed) %>% 
  chuck(1) %>% 
  plotly::ggplotly()
```

### Proportional Barplots

```{r mehg inlet perc barplots, fig.height = 6, fig.width = 8}
loads_plots_mehg %>% pull(barplots_percent) %>% chuck(1)
```

### Boxplots

Red diamond represents the average
```{r mehg inlet boxplots, fig.height = 6, fig.width = 8}
p <- loads_plots_mehg %>% 
  pull(boxplots) %>% 
  chuck(1) %>% 
  plotly::plotly_build()

# Workaround to have plotly not add markers for outliers
p$x$data <- map(p$x$data, remove_markers)
p
```

### Total Load Barplots

```{r mehg total load barplots}
loads_plots_mehg %>%  
  pull(barplots_total) %>% 
  chuck(1) %>% 
  plotly::ggplotly()

loads_plots_mehg %>% pull(barplots_total_prop) %>% chuck(1)
```

### Summary Statistics of each Inlet

**Including** zero load values for the Fremont and Sacramento Weirs when they weren't spilling:
```{r mehg summ stats inlets}
loads_inlet_summ %>% 
  filter(str_detect(Analyte, "^MeHg")) %>% 
  datatable(rownames = FALSE)
```

**Excluding** zero load values for the Fremont and Sacramento Weirs:
```{r mehg summ stats inlets no zero}
loads_inlet_summ_no0 %>% 
  filter(str_detect(Analyte, "^MeHg")) %>% 
  datatable(rownames = FALSE)
```

### % of total inlet load for each inlet

```{r mehg perc load inlets}
loads_inlet_perc_wide %>% 
  filter(str_detect(Analyte, "^MeHg")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(3:7, digits = 1)
```

### Summary Statistics of % loads

**Including** zero percentage values for the Fremont and Sacramento Weirs when they weren't spilling:
```{r mehg summ stats perc inlets}
loads_inlet_perc_summ %>% 
  filter(str_detect(Analyte, "^MeHg")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(4:11, digits = 1)
```

**Excluding** zero percentage values for the Fremont and Sacramento Weirs:
```{r mehg summ stats perc inlets no zero}
loads_inlet_perc_summ_no0 %>% 
  filter(str_detect(Analyte, "^MeHg")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(4:11, digits = 1)
```

### % of each MeHg Fraction

```{r mehg perc frac}
loads_inlet_perc_frac %>% 
  filter(Analyte == "MeHg") %>% 
  select(-Analyte) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(3:4, digits = 1)
```

### Summary Statistics of % MeHg Fractions

```{r mehg summ stats perc frac}
loads_inlet_perc_frac_summ %>% 
  filter(Analyte == "MeHg") %>% 
  pull(summ_stats) %>% 
  chuck(1) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(4:11, digits = 1)
```


## Hg {.tabset .tabset-pills}

### Facet Barplots

```{r hg inlet facet barplots free, fig.height = 9}
loads_plots_hg <- filter(loads_plots, AnalyteGroup == "Hg")

loads_plots_hg %>% pull(barplots_facet_free) %>% chuck(1)
```

```{r hg inlet facet barplots fixed, fig.height = 5.5, fig.width = 8}
loads_barplots_fixed %>% 
  filter(Analyte == "THg- total") %>% 
  pull(barplots_facet_fixed) %>% 
  chuck(1) %>% 
  plotly::ggplotly()
```

### Proportional Barplots

```{r hg inlet perc barplots, fig.height = 6, fig.width = 8}
loads_plots_hg %>% pull(barplots_percent) %>% chuck(1)
```

### Boxplots

Red diamond represents the average
```{r hg inlet boxplots, fig.height = 6, fig.width = 8}
p <- loads_plots_hg %>% 
  pull(boxplots) %>% 
  chuck(1) %>% 
  plotly::plotly_build()

# Workaround to have plotly not add markers for outliers
p$x$data <- map(p$x$data, remove_markers)
p
```

### Total Load Barplots

```{r hg total load barplots}
loads_plots_hg %>%  
  pull(barplots_total) %>% 
  chuck(1) %>% 
  plotly::ggplotly()

loads_plots_hg %>% pull(barplots_total_prop) %>% chuck(1)
```

### Summary Statistics of each Inlet

**Including** zero load values for the Fremont and Sacramento Weirs when they weren't spilling:
```{r hg summ stats inlets}
loads_inlet_summ %>% 
  filter(str_detect(Analyte, "^THg")) %>% 
  datatable(rownames = FALSE)
```

**Excluding** zero load values for the Fremont and Sacramento Weirs:
```{r hg summ stats inlets no zero}
loads_inlet_summ_no0 %>% 
  filter(str_detect(Analyte, "^THg")) %>% 
  datatable(rownames = FALSE)
```

### % of total inlet load for each inlet

```{r hg perc load inlets}
loads_inlet_perc_wide %>% 
  filter(str_detect(Analyte, "^THg")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(3:7, digits = 1)
```

### Summary Statistics of % loads

**Including** zero percentage values for the Fremont and Sacramento Weirs when they weren't spilling:
```{r hg summ stats perc inlets}
loads_inlet_perc_summ %>% 
  filter(str_detect(Analyte, "^THg")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(4:11, digits = 1)
```

**Excluding** zero percentage values for the Fremont and Sacramento Weirs:
```{r hg summ stats perc inlets no zero}
loads_inlet_perc_summ_no0 %>% 
  filter(str_detect(Analyte, "^THg")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(4:11, digits = 1)
```

### % of each Hg Fraction

```{r hg perc frac}
loads_inlet_perc_frac %>% 
  filter(Analyte == "THg") %>% 
  select(-Analyte) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(3:4, digits = 1)
```

### Summary Statistics of % Hg Fractions

```{r hg summ stats perc frac}
loads_inlet_perc_frac_summ %>% 
  filter(Analyte == "THg") %>% 
  pull(summ_stats) %>% 
  chuck(1) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(4:11, digits = 1)
```


## Percent MeHg of Hg loads {.tabset .tabset-pills}

```{r create perc mehg plots}
loads_plots_perc_mehg <- loads_inlet_perc_mehg %>% 
  # Convert per_mehg to a percent and round to the tenths place
  mutate(per_mehg = round(per_mehg * 100, 1)) %>% 
  group_nest(Fraction) %>% 
  mutate(plots = map(data, .f = barplot_perc_mehg))
```

### Filtered Fraction

```{r perc mehg barplot filtered, fig.height = 7, fig.width = 9}
loads_plots_perc_mehg %>% 
  filter(Fraction == "Filtered") %>% 
  pull(plots) %>% 
  chuck(1) %>% 
  plotly::ggplotly()
```

### Particulate Fraction

```{r perc mehg barplot particulate, fig.height = 7, fig.width = 9}
loads_plots_perc_mehg %>% 
  filter(Fraction == "Particulate") %>% 
  pull(plots) %>% 
  chuck(1) %>% 
  plotly::ggplotly()
```

### All Fractions

```{r perc mehg barplot total, fig.height = 7, fig.width = 9}
loads_plots_perc_mehg %>% 
  filter(Fraction == "Total") %>% 
  pull(plots) %>% 
  chuck(1) %>% 
 plotly::ggplotly()
```

### Summary Statistics for Sampling Events in 2017

```{r perc mehg summ stats 2017}
loads_inlet_perc_mehg %>% 
  filter(str_detect(SamplingEvent, "17$")) %>% 
  summ_stat(per_mehg, StationName, Fraction) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(4:11, digits = 1)
```


## Organic Carbon {.tabset .tabset-pills}

### Facet Barplots

```{r oc inlet facet barplots free, fig.height = 9}
loads_plots_oc <- filter(loads_plots, AnalyteGroup == "OrgC")

loads_plots_oc %>% pull(barplots_facet_free) %>% chuck(1)
```

```{r oc inlet facet barplots fixed, fig.height = 5.5, fig.width = 8}
loads_barplots_fixed %>% 
  filter(Analyte == "TOC") %>% 
  pull(barplots_facet_fixed) %>% 
  chuck(1) %>% 
  plotly::ggplotly()
```

### Proportional Barplots

```{r oc inlet perc barplots, fig.height = 6, fig.width = 8}
loads_plots_oc %>% pull(barplots_percent) %>% chuck(1)
```

### Boxplots

Red diamond represents the average
```{r oc inlet boxplots, fig.height = 6, fig.width = 8}
p <- loads_plots_oc %>% 
  pull(boxplots) %>% 
  chuck(1) %>% 
  plotly::plotly_build()

# Workaround to have plotly not add markers for outliers
p$x$data <- map(p$x$data, remove_markers)
p
```

### Total Load Barplots

```{r oc total load barplots}
loads_plots_oc %>%  
  pull(barplots_total) %>% 
  chuck(1) %>% 
  plotly::ggplotly()

loads_plots_oc %>% pull(barplots_total_prop) %>% chuck(1)
```

### Summary Statistics of each Inlet

**Including** zero load values for the Fremont and Sacramento Weirs when they weren't spilling:
```{r oc summ stats inlets}
loads_inlet_summ %>% 
  filter(str_detect(Analyte, "OC$")) %>% 
  datatable(rownames = FALSE)
```

**Excluding** zero load values for the Fremont and Sacramento Weirs:
```{r oc summ stats inlets no zero}
loads_inlet_summ_no0 %>% 
  filter(str_detect(Analyte, "OC$")) %>% 
  datatable(rownames = FALSE)
```

### % of total inlet load for each inlet

```{r oc perc load inlets}
loads_inlet_perc_wide %>% 
  filter(str_detect(Analyte, "OC$")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(3:7, digits = 1)
```

### Summary Statistics of % loads

**Including** zero percentage values for the Fremont and Sacramento Weirs when they weren't spilling:
```{r oc summ stats perc inlets}
loads_inlet_perc_summ %>% 
  filter(str_detect(Analyte, "OC$")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(4:11, digits = 1)
```

**Excluding** zero percentage values for the Fremont and Sacramento Weirs:
```{r oc summ stats perc inlets no zero}
loads_inlet_perc_summ_no0 %>% 
  filter(str_detect(Analyte, "OC$")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(4:11, digits = 1)
```


## Suspended Solids {.tabset .tabset-pills}

### Facet Barplots

```{r ss inlet facet barplots free, fig.height = 9}
loads_plots_ss <- filter(loads_plots, AnalyteGroup == "SuspSolids")

loads_plots_ss %>% pull(barplots_facet_free) %>% chuck(1)
```

```{r ss inlet facet barplots fixed, fig.height = 5.5, fig.width = 8}
loads_barplots_fixed %>% 
  filter(Analyte == "TSS") %>% 
  pull(barplots_facet_fixed) %>% 
  chuck(1) %>% 
  plotly::ggplotly()
```

### Proportional Barplots

```{r ss inlet perc barplots}
loads_plots_ss %>% pull(barplots_percent) %>% chuck(1)
```

### Boxplots

Red diamond represents the average
```{r ss inlet boxplots, fig.height = 6, fig.width = 8}
p <- loads_plots_ss %>% 
  pull(boxplots) %>% 
  chuck(1) %>% 
  plotly::plotly_build()

p$x$data <- map(p$x$data, remove_markers)
p
```

### Total Load Barplots

```{r tss total load barplots}
p <- loads_total_clean_r %>% 
  filter(Analyte == "TSS") %>% 
  ggplot(aes(x = SamplingEvent, y = total_load)) +
  geom_col() +
  labs(
    title = "TSS",
    x = NULL,
    y = "TSS Loads (1,000 kg/day)"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

plotly::ggplotly(p)
```

```{r vss total load barplots}
p <- loads_total_clean_r %>% 
  filter(Analyte == "VSS") %>% 
  ggplot(aes(x = SamplingEvent, y = total_load)) +
  geom_col() +
  labs(
    title = "VSS",
    x = NULL,
    y = "VSS Loads (1,000 kg/day)"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

plotly::ggplotly(p)
```

### Summary Statistics of each Inlet

**Including** zero load values for the Fremont and Sacramento Weirs when they weren't spilling:
```{r ss summ stats inlets}
loads_inlet_summ %>% 
  filter(str_detect(Analyte, "SS$")) %>% 
  datatable(rownames = FALSE)
```

**Excluding** zero load values for the Fremont and Sacramento Weirs:
```{r ss summ stats inlets no zero}
loads_inlet_summ_no0 %>% 
  filter(str_detect(Analyte, "SS$")) %>% 
  datatable(rownames = FALSE)
```

### % of total inlet load for each inlet

```{r ss perc load inlets}
loads_inlet_perc_wide %>% 
  filter(str_detect(Analyte, "SS$")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(3:7, digits = 1)
```

### Summary Statistics of % loads

**Including** zero percentage values for the Fremont and Sacramento Weirs when they weren't spilling:
```{r ss summ stats perc inlets}
loads_inlet_perc_summ %>% 
  filter(str_detect(Analyte, "SS$")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(4:11, digits = 1)
```

**Excluding** zero percentage values for the Fremont and Sacramento Weirs:
```{r ss summ stats perc inlets no zero}
loads_inlet_perc_summ_no0 %>% 
  filter(str_detect(Analyte, "SS$")) %>% 
  datatable(rownames = FALSE) %>% 
  formatPercentage(4:11, digits = 1)
```


# Extra Analyses in Technical Appendix

```{r first flush analysis}
# Calculate the percent reduction of input loads between the first sampling event in 2017 and the largest sampling event in 2017 in terms of inflow (February 14)

loads_total_ff <- loads_total_clean %>% 
  filter(str_detect(SamplingEvent, "Jan 11|Feb 14")) %>% 
  pivot_wider(
    id_cols = Analyte,
    names_from = SamplingEvent,
    values_from = total_load
  ) %>% 
  rename(
    jan11 = "Jan 11-12, 2017",
    feb14 = "Feb 14-15, 2017"
  ) %>% 
  mutate(perc_lower = round((jan11 - feb14)/jan11 * 100))
```

```{r perc load sw 3 events 2017}
# Calculate average load percentages for Sacramento Weir during 3 largest sampling events in 2017

sac_weir17 <- loads_inlet_perc %>% 
  filter(
    SamplingEvent %in% c(
      "Jan 11-12, 2017", 
      "Feb 14-15, 2017", 
      "Mar 1-2, 2017"
    ),
    StationName == "Sacramento Weir"
  ) %>% 
  group_by(Analyte) %>% 
  summarize(avg_perc_load = round(mean(percent_load) * 100))
```

```{r perc umehg load 2017 fw spill}
# Calculate average load percentages of uMeHg for CCSB and KLRC during 8 events in 2017 when the Fremont Weir was spilling

loads_inlet_perc_fw_spill <- loads_inlet_perc %>% 
  filter(
    str_detect(SamplingEvent, "17$"),
    Analyte == "MeHg- total"
  ) %>% 
  pivot_wider(
    id_cols = SamplingEvent,
    names_from = StationName,
    values_from = percent_load
  ) %>% 
  rename(fremont = "Fremont Weir") %>% 
  filter(fremont != 0)

# KLRC average during spill events
klrc_fw_spill <- mean(loads_inlet_perc_fw_spill$KLRC)

# CCSB average during spill events
ccsb_fw_spill <- mean(loads_inlet_perc_fw_spill$CCSB)
```

```{r perc part fraction total inlet}
# Calculate the percentage of the total Hg and MeHg inlet loads in the particulate fraction in 2017

loads_total_avg <- loads_total_clean %>% 
  # Filter load data
  filter(
    str_detect(SamplingEvent, "17$"),
    AnalyteGroup %in% c("Hg", "MeHg")
  ) %>% 
  # Calculate averages for each analyte
  group_by(Analyte) %>% 
  summarize(avg_total_load = mean(total_load))

# Calculate percentage in the particulate fraction
loads_total_part_hg <- loads_total_avg$avg_total_load[5]/(loads_total_avg$avg_total_load[4] + loads_total_avg$avg_total_load[5])

loads_total_part_mehg <- loads_total_avg$avg_total_load[2]/(loads_total_avg$avg_total_load[1] + loads_total_avg$avg_total_load[2])
```


February 14th, approximately a month later, was the sampling event with the largest inflow at 188,700 cfs, but the total inlet loads were not as large as the January 11th event.  This largest inflow event had uHg, uMeHg, and TSS loads that were `r loads_total_ff %>% filter(Analyte == "THg- total") %>% pull(perc_lower)`, `r loads_total_ff %>% filter(Analyte == "MeHg- total") %>% pull(perc_lower)`, and `r loads_total_ff %>% filter(Analyte == "TSS") %>% pull(perc_lower)`% lower than the January 11th event, respectively.

***

During these three sampling events, the Sacramento Weir provided averages of 19% of the total inflow to the Bypass, and `r sac_weir17 %>% filter(Analyte == "THg- total") %>% pull(avg_perc_load)`, `r sac_weir17 %>% filter(Analyte == "MeHg- total") %>% pull(avg_perc_load)`, and `r sac_weir17 %>% filter(Analyte == "TSS") %>% pull(avg_perc_load)`% of the total input load of uHg, uMeHg, and TSS, respectively

***

KLRC contributed an average of `r round(klrc_fw_spill * 100)`% of the uMeHg input load when the Fremont Weir was spilling. CCSB contributed an average of `r round(ccsb_fw_spill * 100)`%.

***

Overall, `r round(loads_total_part_hg * 100)`% and `r round(loads_total_part_mehg * 100)`% of the total inlet loads of Hg and MeHg, respectively, were in the particulate fraction during the 2017 flood

