---
title: "Net Loads and Measurement Uncertainty"
author: "Dave Bosworth"
date: "7/28/2020"
output: 
  html_document: 
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

It's important to know whether the percent difference between a pair of import and export loads is less than their measurement uncertainty. If this is the case, we would ***not*** have confidence that there was a measurable change between the import and export loads. This analysis seeks to answer whether this is the case with any of the net loads reported in the Technical Appendix for the Mass Balance study.

# Methods

The percent differences were calculated by dividing net loads by the average of their associated import and export loads. All load values were used as is and were ***not rounded*** to their proper number of significant figures before the calculation. The upper bound for the measurement uncertainty of the loads provided in the Technical Appendix is +/- 11%. All values less than this criteria were highlighted in the tables below. 

# Data Import and Processing

```{r load packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(openwaterhg)
library(gt)
```

```{r process data and create gt tables, message = FALSE}
# Import total and net load data
source("YB_Mass_Balance/Loads/Import_Total_Load_Data.R")
source("YB_Mass_Balance/Loads/Import_Net_Load_Data.R")

# Combine total and net loads
loads_total_w <- loads_total %>% 
  pivot_wider(
    id_cols = -c(LoadUnits, digits), 
    names_from = LocType,
    values_from = total_load
  ) %>% 
  rename(BLI = "Below Liberty")

loads_net_w <- loads_net %>% 
  pivot_wider(
    id_cols = -c(LoadUnits, digits),
    names_from = Reach,
    values_from = net_load
  )

loads_total_net_c <- left_join(loads_total_w, loads_net_w)

# Calculate percent differences for individual Hg, MeHg, TSS, and OC net loads- 2017 events only
loads_total_net_pd <- loads_total_net_c %>% 
  filter(
    Year == 2017, 
    str_detect(Analyte, "OC$|Hg|SS$")
  ) %>% 
  mutate(
    perc_diff_Upper = round(Upper/((Inlet + Outlet)/2), 3),
    perc_diff_Liberty = round(Liberty/((Outlet + BLI)/2), 3),
    perc_diff_Entire = round(Entire/((Inlet + BLI)/2), 3)
  ) %>%
  # sort by Sampling Event
  conv_fact_samplingevent() %>% 
  arrange(SamplingEvent) %>% 
  # Shorten Analyte names and define order
  mutate(
    Analyte = recode_factor(
      Analyte,
      "THg- total" = "uHg",
      "THg- filtered" = "fHg",
      "THg- particulate" = "pHg",
      "MeHg- total" = "uMeHg",
      "MeHg- filtered" = "fMeHg",
      "MeHg- particulate" = "pMeHg",
      TOC = "TOC",
      DOC = "DOC",
      POC = "POC",
      TSS = "TSS",
      VSS = "VSS"
    )
  )

# Create nested dataframes with summaries for each Analyte and Reach
upper_summ <- loads_total_net_pd %>% 
  select(
    SamplingEvent, 
    Analyte, 
    Inlet, 
    Outlet, 
    contains("Upper") 
  ) %>% 
  rename(
    net_load = Upper,
    perc_diff = perc_diff_Upper
  )

liberty_summ <- loads_total_net_pd %>% 
  select(
    SamplingEvent, 
    Analyte, 
    Outlet, 
    BLI,
    contains("Liberty")
  ) %>% 
  filter(!is.na(BLI)) %>% 
  rename(
    net_load = Liberty,
    perc_diff = perc_diff_Liberty
  )

entire_summ <- loads_total_net_pd %>% 
  select(
    SamplingEvent, 
    Analyte, 
    Inlet, 
    BLI,
    contains("Entire")
  ) %>% 
  filter(!is.na(BLI)) %>% 
  rename(
    net_load = Entire,
    perc_diff = perc_diff_Entire
  )

# Create function to make gt tables of individual percent differences of net loads for each Reach
create_gt_tbl <- function(df, reach_name, param_name, avg_val) {
  # Define measurement uncertainty for cell formatting
  meas_uncert <- 0.11
  
  # Create table
  gt_table <- df %>%
    gt(rowname_col = "SamplingEvent") %>% 
    fmt_number(columns = 2:4) %>% 
    fmt_percent(vars(perc_diff), decimals = 1) %>% 
    tab_style(
      style = list(
        cell_fill(color = "red"),
        cell_text(weight = "bold")
      ),
      locations = cells_body(
        columns = vars(perc_diff),
        rows = abs(perc_diff) <= meas_uncert
      )
    ) %>%  
    tab_header(
      title = paste0("Percent differences of ", param_name, " net loads for the ", reach_name),
      subtitle = "Highlighted cells are less than the measurement uncertainty"
    ) %>% 
    cols_label(
      net_load = "Net Load",
      perc_diff = html("Percent<br>Difference")
    ) %>% 
    grand_summary_rows(
      columns = 2:4,
      fns = list(
        "Average" = "mean",
        "Standard Deviation" = "sd"
      )
    ) %>% 
    grand_summary_rows(
      columns = vars(perc_diff),
      fns = list(
        "Average" = "mean",
        "Standard Deviation" = "sd"
      ),
      formatter = fmt_percent,
      decimals = 1
    )
  
  # Format average percent difference value if less than meas_uncert
  if (abs(avg_val) <= meas_uncert) {
    gt_table <- gt_table %>% 
      tab_style(
        style = list(
          cell_fill(color = "red"),
          cell_text(weight = "bold")
        ),
        locations = cells_grand_summary(
          columns = vars(perc_diff),
          rows = "Average"
        )
      )
  }
  
  return(gt_table)
}

# Combine summary dataframes into a nested tibble and create gt tables for each Analyte and Reach
all_summ <- 
  list(
    "Upper Reach" = upper_summ,
    "Liberty Island Reach" = liberty_summ,
    "Entire Bypass" = entire_summ
  ) %>% 
  map(~group_nest(.x, Analyte)) %>% 
  bind_rows(.id = "Reach") %>% 
  mutate(
    avg_perc_diff = map_dbl(data, ~mean(pull(.x, perc_diff))),
    gt_tbl = pmap(
      list(data, Reach, Analyte, avg_perc_diff),
      .f = create_gt_tbl
    )
  ) %>% 
  select(-c(data, avg_perc_diff)) %>% 
  group_nest(Analyte)
```


# Results {.tabset .tabset-pills}

```{r print gt tables, echo = FALSE, message = FALSE, results = "asis"}
for (i in 1:nrow(all_summ)) {
  # Create heading for each Analyte
  cat("## ", as.character(all_summ$Analyte[i]), " {.tabset .tabset-pills}\n\n")
  for (j in 1:nrow(all_summ$data[[i]])) {
    # Create subheadings for each Reach
    cat("### ", as.character(all_summ$data[[i]]$Reach[j]), "\n\n") 
    # Display table
    print(all_summ$data[[i]]$gt_tbl[[j]])
    cat("\n\n")  
  }
}
```


