---
title: "Net Loads Analysis template"
author: "Dave Bosworth"
date: "4/18/2020"
output: html_document
---

## {{parameter}} {.tabset .tabset-pills}

```{r filter {{parameter}}}
filt_obj <- all_obj %>% filter(AnalyteGroup == "{{parameter}}")
```

### Total Load Barplots

```{r {{parameter}} total load barplots, fig.height = 6}
filt_obj %>% pull(t_barplots) %>% chuck(1)
```

### Total Load Boxplots

Red diamond represents the average
```{r {{parameter}} total load boxplots, fig.height = 5.5}
p <- filt_obj %>% 
  pull(t_boxplots) %>% 
  chuck(1) %>% 
  plotly::plotly_build()

# Workaround to have plotly not add markers for outliers
p$x$data <- map(p$x$data, remove_markers)
p
```

### Total Load vs. Inflow Scatterplots

```{r {{parameter}} total load vs flow plots, fig.height = 7, fig.width = 7}
filt_obj %>% pull(t_scatterplots) %>% chuck(1)
```

### Total Load Data Tables

```{r unnest {{parameter}} total load data tables}
filt_obj_unnest <- filt_obj %>% 
  pull(t_loads_table) %>% 
  chuck(1)
```

```{r {{parameter}} total load data tables, echo = FALSE, message = FALSE, results = "asis"}
for (i in 1:nrow(filt_obj_unnest)) {
  # Create subheadings for each LocType
  cat("#### ", as.character(filt_obj_unnest$LocType[i]), "\n\n") 
  # Display table of total loads
  print(filt_obj_unnest$load_table[[i]])
  cat("\n\n")
}
```

### Total Load Summary Statistics

```{r {{parameter}} total load summ stats}
filt_obj %>% pull(t_loads_summ) %>% chuck(1)
```

### Net Load Barplots

Free y-axis scale:
```{r {{parameter}} net load barplots free, fig.height = 6}
filt_obj %>% pull(n_barplots_free) %>% chuck(1)
```

Fixed y-axis scale:
```{r {{parameter}} net load barplots fixed, fig.height = 7}
filt_obj %>% pull(n_barplots_fixed) %>% chuck(1)
```

### Net Load Boxplots

Red diamond represents the average
```{r {{parameter}} net load boxplots, fig.height = 5.5}
p <- filt_obj %>% 
  pull(n_boxplots) %>% 
  chuck(1) %>% 
  plotly::plotly_build()

# Workaround to have plotly not add markers for outliers
p$x$data <- map(p$x$data, remove_markers)
p
```

### Net Load vs. Inflow Scatterplots

```{r unnest {{parameter}} net load vs flow plots}
filt_obj_unnest <- filt_obj %>% 
  pull(n_scatterplots) %>% 
  chuck(1)
```

```{r {{parameter}} net load vs flow plots, echo = FALSE, message = FALSE, results = "asis", fig.width = 9, fig.height = 4}
for (i in 1:nrow(filt_obj_unnest)) {
  # Create subheadings for each Reach
  cat("#### ", as.character(filt_obj_unnest$Reach[i]), "\n\n") 
  # Display grouped plot
  print(filt_obj_unnest$grouped_plots[[i]])
  cat("\n\n")
}
```

### Net Load Data Tables

```{r unnest {{parameter}} net load data tables}
filt_obj_unnest <- filt_obj %>% 
  pull(n_loads_table) %>% 
  chuck(1)
```

```{r {{parameter}} net load data tables, echo = FALSE, message = FALSE, results = "asis"}
for (i in 1:nrow(filt_obj_unnest)) {
  # Create subheadings for each Reach
  cat("#### ", as.character(filt_obj_unnest$Reach[i]), "\n\n") 
  # Display table of net loads
  print(filt_obj_unnest$load_table[[i]])
  cat("\n\n")
}
```

### Net Load Summary Statistics

```{r {{parameter}} net load summ stats}
filt_obj %>% pull(n_loads_summ) %>% chuck(1)
```

