---
title: "Net Loads for the Yolo Bypass"
author: "Dave Bosworth"
date: "4/1/2020"
output: 
  html_document: 
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float:
      collapsed: no
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import and Prepare Data

Load R packages necessary for this analysis:
```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(scales)
library(rlang)
library(openwaterhg)
library(DT)
```

Calculate total input and output loads for each sampling event and parameter:
```{r calc total loads}
total_loads <- loads_calc %>% 
  # Only include a subset of the data
  filter(str_detect(Analyte, "OC$|Hg|SS$")) %>%
  # Add an AnalyteGroup variable
  mutate(
    AnalyteGroup = case_when(
      str_detect(Analyte, "^MeHg") ~ "MeHg",
      str_detect(Analyte, "^THg") ~ "Hg",
      str_detect(Analyte, "OC$") ~ "OrgC",
      str_detect(Analyte, "SS$") ~ "SuspSolids"
    )
  ) %>% 
  # Group and sum load data
  group_by(SamplingEvent, Analyte, LoadUnits, AnalyteGroup, LocType) %>% 
  summarize(total_load = sum(Load)) %>% 
  ungroup() %>% 
  # Apply plot order
  conv_fact_samplingevent() %>% 
  mutate(LocType = factor(LocType, levels = c("Inlet", "Outlet", "Below Liberty")))
```

Calculate net loads for each reach (Upper, Liberty Island, Entire Bypass) grouped by sampling event and parameter:
```{r calc net loads}
net_loads <- total_loads %>% 
  pivot_wider(names_from = LocType, values_from = total_load) %>% 
  rename(below_liberty = "Below Liberty") %>% 
  # Calculate net loads for each reach
  mutate(
    Upper = Outlet - Inlet,
    Liberty = below_liberty - Outlet,
    Entire = below_liberty - Inlet
  ) %>% 
  select(-c(Inlet:below_liberty)) %>% 
  pivot_longer(
    cols = Upper:Entire,
    names_to = "Reach",
    values_to = "net_load"
  ) %>% 
  filter(!is.na(net_load)) %>% 
  # Apply plot order
  mutate(Reach = factor(Reach, levels = c("Upper", "Liberty", "Entire")))
```

Add total inlet flows to total_loads and net_loads dataframes:
```{r add flow data, message = FALSE}
# Prepare flow data df to join with loads df's
total_inflows <- daily_flow_data_se %>% 
  # Only include inlet flows
  filter(LocType == "Inlet") %>% 
  # Group by and sum flow data
  group_by(SamplingEvent) %>% 
  summarize(total_inflow = sum(Flow)) %>% 
  ungroup() %>% 
  # Apply plot order
  conv_fact_samplingevent()

# Join inlet flows to total_loads df
total_loads <- left_join(total_loads, total_inflows)

# Join inlet flows to net_loads df
net_loads <- left_join(net_loads, total_inflows)
```

# Run Calculations

Calculate summary statistics of total loads:
```{r calc summ stats total loads}
total_loads_summ <- total_loads %>% 
  summ_stat(total_load, LocType, Analyte)
```

Calculate summary statistics of net loads:
```{r calc summ stats net loads}
net_loads_summ <- net_loads %>% 
  summ_stat(net_load, Reach, Analyte)
```

# Create Plot Functions

Function for barplots for total and net loads:
```{r fun barplot loads}
barplot_load <- function(df, y_var, fill_var, type = c("total", "net")) {
  # evaluate choices for type
  type <- match.arg(type, c("total", "net"))
  
  # convert variables to enquo
  y_var_enquo <- enquo(y_var)
  fill_var_enquo <- enquo(fill_var)
  
  # create plot
  p <- 
    ggplot(
      data = df,
      aes(
        x = SamplingEvent, 
        y = !!y_var_enquo, 
        fill = !!fill_var_enquo
      )
    ) +
    geom_col(position = "dodge") +
    facet_wrap(
      vars(Analyte),
      scales = "free_y"
    ) +
    add_gen_color_pal(3, "fill") +
    theme_owhg(x_axis_v = TRUE) +
    theme(legend.position = "bottom")
  
  # define y-axis labels based on type arg
  if (type == "total") {
    p <- p +
      labs(
        title = NULL,
        x = NULL,
        y = paste0("Total Loads (", df$LoadUnits[1], ")")
      )
  } else {
    p <- p +
      labs(
        title = NULL,
        x = NULL,
        y = paste0("Net Loads (", df$LoadUnits[1], ")")
      )
  }
  
  return(p)
}
```

Function for boxplots for total and net loads:
```{r fun boxplot loads}
boxplot_load <- function(df, x_var, y_var, type = c("total", "net")) {
  # evaluate choices for type
  type <- match.arg(type, c("total", "net"))
  
  # convert variables to enquo
  x_var_enquo <- enquo(x_var)
  y_var_enquo <- enquo(y_var)
  
  # create plot
  p <- 
    ggplot(
      data = df,
      aes(
        x = !!x_var_enquo,
        y = !!y_var_enquo
      )
    ) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(width = 0.25) +
    stat_summary( 
      fun = "mean", #add a symbol representing the mean of each group to the plot
      color = "red",
      geom = "point",
      shape = 9, #diamond cross shape
      size = 2 
    ) +
    facet_wrap(vars(Analyte)) +
    theme_owhg(x_axis_v = TRUE)
  
  # define title labels based on type arg
  if (type == "total") {
    p <- p +
      labs(
        title = paste0("Total Loads (", df$LoadUnits[1], ")"),
        x = NULL,
        y = NULL
      )
  } else {
    p <- p +
      labs(
        title = paste0("Net Loads (", df$LoadUnits[1], ")"),
        x = NULL,
        y = NULL
      )
  }
    
  return(p)
}
```

Function for total loads vs. total inflow scatterplots:
```{r fun total load vs inflow scatterplots}
scattplot_tload_flow <- function(df) {
  p <- 
    ggplot(
      data = df,
      aes(
        x = total_inflow, 
        y = total_load, 
        color = LocType
      )
    ) +
    geom_point(size = 1.8, alpha = 0.7) +
    facet_wrap(
      vars(Analyte),
      scales = "free_y"
    ) +
    labs(
        title = NULL,
        x = "Total Inflow (cfs)",
        y = paste0("Total Load (", df$LoadUnits[1], ")")
    ) +
    scale_x_continuous(labels = label_comma()) +
    add_gen_color_pal(3, "color") +
    theme_owhg() +
    theme(legend.position = "bottom")
  
  return(p)
}
```

# Create Plots

Total Loads:
```{r create plots total loads}
total_loads_plots <- total_loads %>% 
  group_nest(AnalyteGroup) %>% 
  mutate(
    barplots = map(
      data,
      .f = barplot_load,
      y_var = total_load,
      fill_var = LocType,
      type = "total"
    ), 
    boxplots = map(
      data,
      .f = boxplot_load,
      x_var = LocType,
      y_var = total_load,
      type = "total"
    ), 
    scatterplots = map(data, .f = scattplot_tload_flow)
  )
```

Net Loads:
```{r create plots net loads}
net_loads_plots <- net_loads %>% 
  group_nest(AnalyteGroup) %>% 
  mutate(
    barplots = map(
      data,
      .f = barplot_load,
      y_var = net_load,
      fill_var = Reach,
      type = "net"
    ), 
    boxplots = map(
      data,
      .f = boxplot_load,
      x_var = Reach,
      y_var = net_load,
      type = "net"
    )
  )
```

# Plots and Summary Statistics

## MeHg

### Total Load Barplots

```{r mehg total load barplots, fig.height = 6, fig.width = 8}
total_loads_plots_mehg <- filter(total_loads_plots, AnalyteGroup == "MeHg")

total_loads_plots_mehg %>% pull(barplots) %>% chuck(1)
```

### Total Load Boxplots

```{r mehg total load boxplots, fig.height = 5.5}
total_loads_plots_mehg %>% pull(boxplots) %>% chuck(1) %>% plotly::ggplotly()
```


### Total Load vs. Inflow Scatterplots

```{r mehg load vs flow plots, fig.height = 4.5, fig.width = 8}
total_loads_plots_mehg %>% pull(scatterplots) %>% chuck(1)
```

### Total Load Summary Statistics

```{r mehg total load summ stats}
total_loads_summ %>% 
  filter(str_detect(Analyte, "^MeHg")) %>% 
  datatable(rownames = FALSE) %>% 
  formatRound(4:11, digits = 2)
```

### Net Load Barplots

### Net Load Boxplots

### Net Load Summary Statistics


## Hg

### Total Loads
</br>

#### Barplots

```{r hg total load barplots, fig.height = 6, fig.width = 8}
total_loads_plots_hg <- filter(total_loads_plots, AnalyteGroup == "Hg")

total_loads_plots_hg %>% pull(barplots) %>% chuck(1)
```

#### Boxplots

```{r hg total load boxplots, fig.height = 5.5}
total_loads_plots_hg %>% pull(boxplots) %>% chuck(1) %>% plotly::ggplotly()
```

#### Load vs. Inflow Scatterplots

```{r hg load vs flow plots, fig.height = 4.5, fig.width = 8}
total_loads_plots_hg %>% pull(scatterplots) %>% chuck(1)
```

#### Summary Statistics

```{r hg total load summ stats}
total_loads_summ %>% 
  filter(str_detect(Analyte, "^THg")) %>% 
  datatable(rownames = FALSE) %>% 
  formatRound(4:11, digits = 1)
```

## Organic Carbon

### Total Loads
</br>

#### Barplots

```{r oc total load barplots, fig.height = 6, fig.width = 8}
total_loads_plots_oc <- filter(total_loads_plots, AnalyteGroup == "OrgC")

total_loads_plots_oc %>% pull(barplots) %>% chuck(1)
```

#### Boxplots

```{r oc total load boxplots, fig.height = 5.5}
total_loads_plots_oc %>% pull(boxplots) %>% chuck(1) %>% plotly::ggplotly()
```

#### Load vs. Inflow Scatterplots

```{r oc load vs flow plots, fig.height = 4.5, fig.width = 8}
total_loads_plots_oc %>% pull(scatterplots) %>% chuck(1)
```

#### Summary Statistics

```{r oc total load summ stats}
total_loads_summ %>% 
  filter(str_detect(Analyte, "OC$")) %>% 
  datatable(rownames = FALSE) %>% 
  formatRound(4:11, digits = 1)
```

## Suspended Solids

### Total Loads
</br>

#### Barplots

```{r ss total load barplots, fig.height = 6, fig.width = 8}
total_loads_plots_ss <- filter(total_loads_plots, AnalyteGroup == "SuspSolids")

total_loads_plots_ss %>% pull(barplots) %>% chuck(1)
```

#### Boxplots

```{r ss total load boxplots, fig.height = 5.5}
total_loads_plots_ss %>% pull(boxplots) %>% chuck(1) %>% plotly::ggplotly()
```

#### Load vs. Inflow Scatterplots

```{r ss load vs flow plots, fig.height = 4.5, fig.width = 8}
total_loads_plots_ss %>% pull(scatterplots) %>% chuck(1)
```

#### Summary Statistics

```{r ss total load summ stats}
total_loads_summ %>% 
  filter(str_detect(Analyte, "SS$")) %>% 
  datatable(rownames = FALSE) %>% 
  formatRound(4:11, digits = 0)
```

