---
title: "Net Loads for the Yolo Bypass"
author: "Dave Bosworth"
date: "4/1/2020"
output: 
  html_document: 
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import and Prepare Data

Load R packages necessary for this analysis:
```{r load packages, message = FALSE, warning = FALSE}
# Load packages
library(tidyverse)
library(scales)
library(rlang)
library(openwaterhg)
library(knitr)
library(kableExtra)
library(DT)
library(broom)
library(patchwork)
```

Bring in total input and output loads for each sampling event and parameter:
```{r calc total loads}
# Bring in total load data
setwd("C:/Repositories/Open-Water-Hg-Studies/YB_Mass_Balance/Loads")
source("Import_Total_Load_Data.R")

# Prepare total load data for plots and summaries
loads_total_clean <- loads_total %>% 
  # Only include a subset of the data
  filter(str_detect(Analyte, "OC$|Hg|SS$")) %>%
  # Add an AnalyteGroup variable
  mutate(
    AnalyteGroup = case_when(
      str_detect(Analyte, "^MeHg") ~ "MeHg",
      str_detect(Analyte, "^THg") ~ "Hg",
      str_detect(Analyte, "OC$") ~ "Organic Carbon",
      str_detect(Analyte, "SS$") ~ "Suspended Solids"
    )
  ) %>% 
  # Apply plot order
  conv_fact_samplingevent() %>% 
  mutate(LocType = factor(LocType, levels = c("Inlet", "Outlet", "Below Liberty")))

# Clean up
rm(loads_total)
```

Bring in net loads for each reach (Upper, Liberty Island, Entire Bypass) grouped by sampling event and parameter:
```{r calc net loads, message = FALSE}
# Bring in total load data
source("Import_Net_Load_Data.R")

# Prepare net load data for plots and summaries
loads_net_clean <- loads_net %>% 
  # Only include a subset of the data
  filter(str_detect(Analyte, "OC$|Hg|SS$")) %>%
  # Add an AnalyteGroup variable
  mutate(
    AnalyteGroup = case_when(
      str_detect(Analyte, "^MeHg") ~ "MeHg",
      str_detect(Analyte, "^THg") ~ "Hg",
      str_detect(Analyte, "OC$") ~ "Organic Carbon",
      str_detect(Analyte, "SS$") ~ "Suspended Solids"
    )
  ) %>% 
  # Apply plot order
  conv_fact_samplingevent() %>% 
  mutate(Reach = factor(Reach, levels = c("Upper", "Liberty", "Entire")))

# Clean up
rm(loads_net)
```

Add total inlet flows to total_loads and net_loads dataframes:
```{r add flow data, message = FALSE}
# Prepare flow data df to join with loads df's
total_inflows <- daily_flow_data_se %>% 
  # Only include inlet flows
  filter(LocType == "Inlet") %>% 
  # Group by and sum flow data
  group_by(SamplingEvent) %>% 
  summarize(total_inflow = sum(Flow)) %>% 
  ungroup() %>% 
  # Apply plot order
  conv_fact_samplingevent()

# Join inlet flows to total_loads df
loads_total_clean <- left_join(loads_total_clean, total_inflows)

# Join inlet flows to net_loads df
loads_net_clean <- left_join(loads_net_clean, total_inflows)
```

Create two new dataframes with total and net loads only for 2017 sampling events. This will be used for the summary statistics, boxplots, and net load vs flow plots.
```{r filter 2017 data}
loads_total_clean17 <- loads_total_clean %>% filter(str_detect(SamplingEvent, "17$"))

loads_net_clean17 <- loads_net_clean %>% filter(str_detect(SamplingEvent, "17$"))
```

Create rounded versions of the total and net load dataframes for some of the analyses:
```{r round load data}
loads_total_clean_r <- loads_total_clean %>% 
  mutate(total_load = signif(total_load, digits)) %>% 
  select(-digits)

loads_total_clean17_r <- loads_total_clean17 %>% 
  mutate(total_load = signif(total_load, digits)) %>% 
  select(-digits)

loads_net_clean_r <- loads_net_clean %>% 
  mutate(net_load = signif(net_load, digits)) %>% 
  select(-digits)

loads_net_clean17_r <- loads_net_clean17 %>% 
  mutate(net_load = signif(net_load, digits)) %>% 
  select(-digits)
```

Create restructured total_loads dataframes grouped by LocType and Analyte Group to display as datatables:
```{r total loads restructure}
loads_total_table <- loads_total_clean_r %>% 
  select(-total_inflow) %>% 
  arrange(SamplingEvent) %>%
  group_nest(LocType, AnalyteGroup) %>% 
  mutate(
    load_table = map(
      data,
      .f = ~pivot_wider(
        .x,
        id_cols = SamplingEvent, 
        names_from = Analyte, 
        values_from = total_load
      ) %>% 
      kable(
        format = "html",
        col.names = gsub("gE", "g E", names(.)),
        format.args = list(
          big.mark = ",", 
          drop0trailing = TRUE
        )
      ) %>% 
      kable_styling(
        "striped", 
        full_width = FALSE, 
        position = "left"
      )
    )
  ) %>% 
  select(-data) %>% 
  group_nest(AnalyteGroup) %>% 
  rename(t_loads_table = data)
```

Create restructured net_loads dataframes grouped by Reach and Analyte Group to display as datatables:
```{r net loads restructure}
loads_net_table <- loads_net_clean_r %>% 
  select(-total_inflow) %>% 
  arrange(SamplingEvent) %>% 
  # Apply conditional formatting to net loads
  mutate(
    net_load = cell_spec(
      format(
        net_load,
        big.mark = ",", 
        scientific = FALSE,
        drop0trailing = TRUE
      ),
      "html",
      color = if_else(net_load < 0, "red", "black")
    )
  ) %>% 
  group_nest(Reach, AnalyteGroup) %>% 
  mutate(
    load_table = map(
      data,
      .f = ~pivot_wider(
        .x,
        id_cols = SamplingEvent, 
        names_from = Analyte, 
        values_from = net_load
      ) %>% 
      kable(
        format = "html",
        col.names = gsub("gE", "g E", names(.)),
        escape = FALSE
      ) %>% 
      kable_styling(
        "striped", 
        full_width = FALSE, 
        position = "left"
      )
    )
  ) %>% 
  select(-data) %>% 
  group_nest(AnalyteGroup) %>% 
  rename(n_loads_table = data)
```


# Run Calculations

Calculate summary statistics of total loads:
```{r calc summ stats total loads, message = FALSE}
# Calculate averages and standard deviations on non-rounded data
loads_total_summ1 <- loads_total_clean17 %>% 
  group_nest(AnalyteGroup) %>% 
  mutate(
    t_loads_summ1 = map(
      data,
      .f = ~group_by(.x, LocType, Analyte) %>% 
        summarize(
          sign_digits = min(digits),
          Mean = signif(mean(total_load), sign_digits),
          StDev = signif(sd(total_load), sign_digits)
        ) %>% 
        ungroup() %>% 
        select(-sign_digits)
    )
  ) %>% 
  select(-data)
  
# Calculate all other summary statistics on rounded data
loads_total_summ2 <- loads_total_clean17_r %>% 
  group_nest(AnalyteGroup) %>% 
  mutate(
    t_loads_summ2 = map(
      data,
      .f = ~summ_stat(.x, total_load, LocType, Analyte) %>% 
        select(-c(Mean, StDev))
    )
  ) %>% 
  select(-data)

# Join Summary Statistics together
loads_total_summ <- 
  left_join(loads_total_summ1, loads_total_summ2) %>% 
  mutate(
    t_loads_summ = map2(
      t_loads_summ2,
      t_loads_summ1,
      .f = ~left_join(.x, .y) %>% 
        select(LocType:N, Mean, StDev, Minimum:IQR) %>% 
        datatable(rownames = FALSE)
    )
  ) %>% 
  select(-c(t_loads_summ1, t_loads_summ2))

# Clean up
rm(loads_total_summ1, loads_total_summ2)
```

Calculate summary statistics of net loads:
```{r calc summ stats net loads, message = FALSE}
# Calculate averages and standard deviations on non-rounded data
loads_net_summ1 <- loads_net_clean17 %>% 
  group_nest(AnalyteGroup) %>% 
  mutate(
    n_loads_summ1 = map(
      data,
      .f = ~group_by(.x, Reach, Analyte) %>% 
        summarize(
          sign_digits = min(digits),
          Mean = signif(mean(net_load), sign_digits),
          StDev = signif(sd(net_load), sign_digits)
        ) %>% 
        ungroup() %>% 
        select(-sign_digits)
    )
  ) %>% 
  select(-data)
  
# Calculate all other summary statistics on rounded data
loads_net_summ2 <- loads_net_clean17_r %>% 
  group_nest(AnalyteGroup) %>% 
  mutate(
    n_loads_summ2 = map(
      data,
      .f = ~summ_stat(.x, net_load, Reach, Analyte) %>% 
        select(-c(Mean, StDev))
    )
  ) %>% 
  select(-data)

# Join Summary Statistics together
loads_net_summ <- 
  left_join(loads_net_summ1, loads_net_summ2) %>% 
  mutate(
    n_loads_summ = map2(
      n_loads_summ2,
      n_loads_summ1,
      .f = ~left_join(.x, .y) %>% 
        select(Reach:N, Mean, StDev, Minimum:IQR) %>% 
        datatable(rownames = FALSE) %>% 
        formatStyle(
          4:11,
          color = styleInterval(0, values = c("red", "black"))
        )
    )
  ) %>% 
  select(-c(n_loads_summ1, n_loads_summ2))

# Clean up
rm(loads_net_summ1, loads_net_summ2)
```


# Create Plot Functions

Function for barplots for total and net loads, free y-axis scale:
```{r fun barplot loads free}
barplot_load_free <- function(df, y_var, fill_var, type = c("total", "net")) {
  # evaluate choices for type
  type <- match.arg(type, c("total", "net"))
  
  # convert variables to enquo
  y_var_enquo <- enquo(y_var)
  fill_var_enquo <- enquo(fill_var)
  
  # create plot
  p <- 
    ggplot(
      data = df,
      aes(
        x = SamplingEvent, 
        y = !!y_var_enquo, 
        fill = !!fill_var_enquo
      )
    ) +
    geom_col(position = "dodge") +
    facet_grid(
      rows= vars(Analyte),
      scales = "free"
    ) +
    add_gen_color_pal(3, "fill") +
    theme_owhg(x_axis_v = TRUE)
  
  # define y-axis labels based on type arg
  if (type == "total") {
    p <- p +
      labs(
        title = NULL,
        x = NULL,
        y = paste0("Total Loads (", df$LoadUnits[1], ")")
      )
  } else {
    p <- p +
      labs(
        title = NULL,
        x = NULL,
        y = paste0("Net Loads (", df$LoadUnits[1], ")")
      )
  }
  
  return(p)
}
```

Function for barplots for net loads, fixed y-axis scale:
```{r fun barplot net loads fixed}
barplot_load_fixed <- function(df) {
   p <- 
    ggplot(
      data = df,
      aes(
        x = SamplingEvent, 
        y = net_load, 
        fill = Reach
      )
    ) +
    geom_col(position = "dodge") +
    facet_grid(rows= vars(Analyte)) +
    labs(
      title = NULL,
      x = NULL,
      y = paste0("Net Loads (", df$LoadUnits[1], ")")
    ) +
    add_gen_color_pal(3, "fill") +
    theme_owhg(x_axis_v = TRUE)
  
  return(p)
}
```

Function for boxplots for total and net loads:
```{r fun boxplot loads}
boxplot_load <- function(df, x_var, y_var, type = c("total", "net")) {
  # evaluate choices for type
  type <- match.arg(type, c("total", "net"))
  
  # convert variables to enquo
  x_var_enquo <- enquo(x_var)
  y_var_enquo <- enquo(y_var)
  
  # create plot
  p <- 
    ggplot(
      data = df,
      aes(
        x = !!x_var_enquo,
        y = !!y_var_enquo
      )
    ) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(width = 0.25) +
    stat_summary( 
      fun = "mean", #add a symbol representing the mean of each group to the plot
      color = "red",
      geom = "point",
      shape = 9, #diamond cross shape
      size = 2 
    ) +
    facet_wrap(vars(Analyte)) +
    theme_owhg(x_axis_v = TRUE)
  
  # define title labels based on type arg
  if (type == "total") {
    p <- p +
      labs(
        title = paste0("Total Loads (", df$LoadUnits[1], ")"),
        x = NULL,
        y = NULL
      )
  } else {
    p <- p +
      labs(
        title = paste0("Net Loads (", df$LoadUnits[1], ")"),
        x = NULL,
        y = NULL
      )
  }
    
  return(p)
}
```

Function for total loads vs. total inflow scatterplots:
```{r fun total load vs inflow scatterplots}
scattplot_tload_flow <- function(df) {
  p <- 
    ggplot(
      data = df,
      aes(
        x = total_inflow, 
        y = total_load
      )
    ) +
    geom_point() +
    facet_grid(
      cols = vars(LocType),
      rows = vars(Analyte),
      scales = "free"
    ) +
    labs(
        title = NULL,
        x = "Total Inflow (cfs)",
        y = paste0("Total Load (", df$LoadUnits[1], ")")
    ) +
    scale_x_continuous(labels = label_comma()) +
    theme_owhg()
  
  return(p)
}
```

Function for net loads vs. total inflow scatterplots:
```{r fun net load vs inflow scatterplots}
scattplot_nload_flow <- function(df, param, rsq, pval) {
  p <- 
    ggplot(
      data = df,
      aes(
        x = total_inflow, 
        y = net_load
      )
    ) +
    geom_point() +
    geom_smooth(
      method = "lm",
      formula = y ~ x,
      se = FALSE
    ) +
    labs(
        title = param,
        subtitle = paste0("R Squared = ", rsq, "%\np-value = ", pval),
        x = "Total Inflow (cfs)",
        y = paste0("Net Load (", df$LoadUnits[1], ")")
    ) +
    scale_x_continuous(labels = label_comma()) +
    scale_y_continuous(labels = label_comma()) +
    theme_owhg()
  
  return(p)
}
```


# Create Plots

Total Loads:
```{r create plots total loads, message = FALSE}
# Barplots and scatterplots- all sampling events, rounded data
loads_total_plots <- loads_total_clean_r %>% 
  group_nest(AnalyteGroup) %>% 
  mutate(
    t_barplots = map(
      data,
      .f = barplot_load_free,
      y_var = total_load,
      fill_var = LocType,
      type = "total"
    ), 
    t_scatterplots = map(data, .f = scattplot_tload_flow)
  ) %>% 
  select(-data)

# Boxplots- just 2017 sampling events, rounded data
loads_total_boxplots <- loads_total_clean17_r %>% 
  group_nest(AnalyteGroup) %>% 
  mutate(
    t_boxplots = map(
      data,
      .f = boxplot_load,
      x_var = LocType,
      y_var = total_load,
      type = "total"
    )
  ) %>% 
  select(-data)
```

Net Loads:
```{r create plots net loads, message = FALSE}
# Barplots- all sampling events, rounded data
loads_net_barplots <- loads_net_clean_r %>% 
  group_nest(AnalyteGroup) %>% 
  mutate(
    n_barplots_free = map(
      data,
      .f = barplot_load_free,
      y_var = net_load,
      fill_var = Reach,
      type = "net"
    ),
    n_barplots_fixed = map(data, .f = barplot_load_fixed)
  ) %>% 
  select(-data)

# Boxplots- just 2017 sampling events, rounded data
loads_net_boxplots <- loads_net_clean17_r %>% 
  group_nest(AnalyteGroup) %>% 
  mutate(
    n_boxplots = map(
      data,
      .f = boxplot_load,
      x_var = Reach,
      y_var = net_load,
      type = "net"
    )
  ) %>% 
  select(-data)

# Net load vs inflow scatterplots- just 2017 sampling events, unrounded data
loads_net_scatterplots <- loads_net_clean17 %>% 
  group_nest(Reach, AnalyteGroup, Analyte) %>% 
  # Run regression analysis and create plots for each Analyte-Reach combination
  mutate(
    model = map(data, ~summary(lm(net_load ~ total_inflow, data = .x))),
    r2 = signif(map_dbl(model, ~glance(.x)$r.squared * 100), 3),
    p_value = signif(map_dbl(model, ~glance(.x)$p.value), 2),
    plot = pmap(
      list(data, Analyte, r2, p_value),
      .f = scattplot_nload_flow
    )
  ) %>% 
  select(Reach, AnalyteGroup, plot) %>% 
  group_nest(Reach, AnalyteGroup) %>% 
  # Create grouped plots by AnalyteGroup for each Reach using patchwork package
  mutate(
    grouped_plots = map(
      data, 
      .f = ~pull(.x, plot) %>% 
        wrap_plots()
    )
  ) %>% 
  select(-data) %>% 
  # Group nest by AnalyteGroup to join with other dataframes
  group_nest(AnalyteGroup) %>% 
  rename(n_scatterplots = data)
```


# Plots and Summary Statistics {.tabset .tabset-pills}

The summary statistics and boxplots for the total and net loads only include the sampling events conducted in 2017. The net load vs inflow scatterplots also include only the 2017 data. The remaining plots include all sampling events conducted throughout the study period (2014-2017).

Join all plots, data tables, and summary statistics together for more efficient code:
```{r join plots tables summ stats, message = FALSE}
all_obj <- 
  reduce(
    list(
      loads_total_plots,
      loads_total_boxplots,
      loads_total_table,
      loads_total_summ,
      loads_net_barplots,
      loads_net_boxplots,
      loads_net_scatterplots,
      loads_net_table,
      loads_net_summ
    ),
    left_join
  ) 

# Clean up
rm(
  loads_total_plots,
  loads_total_boxplots,
  loads_total_table,
  loads_total_summ,
  loads_net_barplots,
  loads_net_boxplots,
  loads_net_scatterplots,
  loads_net_table,
  loads_net_summ
)
```

Create function to have plotly not add markers for outliers:
```{r fun remove plotly markers}
remove_markers <- function(x) {
  if (x$type == "box") {
      x$marker = list(opacity = 0)
    }
    return(x)
}
```

```{r, include = FALSE}
run_analyses <- 
  map(
    c("MeHg", "Hg", "Organic Carbon", "Suspended Solids"),
    .f = ~knit_expand(file = "Net_loads_Analysis_template.Rmd", parameter = .x)
  )
```

`r knit(text = unlist(run_analyses))`

# Extra Analyses in Technical Appendix

```{r inlet vs internal mehg source for entire yb}
mehg_total_entire_yb <- loads_total_clean17 %>% 
  filter(
    !str_detect(SamplingEvent, "Apr 11"),
    Analyte == "MeHg- total"
  ) %>%
  select(-c(LoadUnits, AnalyteGroup)) %>% 
  group_by(LocType) %>% 
  summarize(avg_total_load = mean(total_load))

mehg_inlet <- mehg_total_entire_yb$avg_total_load[1]
mehg_outlet <- mehg_total_entire_yb$avg_total_load[2]
mehg_bli <- mehg_total_entire_yb$avg_total_load[3]
perc_inlet <- round(mehg_inlet/mehg_bli * 100)

```

On average, the inlets to the Bypass accounted for `r perc_inlet`% of this uMeHg export load and net internal production provided the remaining `r 100 - perc_inlet`% or 18 ± 19 g/day (Table 13).  Most of this net internal production of uMeHg occurred in the Upper reach between the inlets and the Stairsteps supplying `r round((mehg_outlet - mehg_inlet)/(mehg_bli - mehg_inlet) * 100)`% of the net load.

***

```{r perc fhg export entire bypass}
hg_total_entire_yb <- loads_total_clean17 %>% 
  filter(
    LocType == "Below Liberty",
    AnalyteGroup == "Hg"
  ) %>% 
  group_by(Analyte) %>% 
  summarize(avg_total_load = mean(total_load))

perc_fhg_total_entire_yb <- 
  hg_total_entire_yb$avg_total_load[1]/(hg_total_entire_yb$avg_total_load[1] + hg_total_entire_yb$avg_total_load[2])
```

The filter-passing mercury fraction was a small component of the total uHg leaving the Bypass to the downstream Delta at `r round(perc_fhg_total_entire_yb * 100)`% of the average uHg load 

***

```{r perc pmehg export upper bypass}
mehg_total_upper_yb <- loads_total_clean17 %>% 
  filter(
    LocType == "Outlet",
    AnalyteGroup == "MeHg"
  ) %>% 
  group_by(Analyte) %>% 
  summarize(avg_total_load = mean(total_load))

perc_pmehg_total_upper_yb <- 
  mehg_total_upper_yb$avg_total_load[2]/(mehg_total_upper_yb$avg_total_load[1] + mehg_total_upper_yb$avg_total_load[2])
```

Overall, `r round(perc_pmehg_total_upper_yb * 100)`% of the MeHg exported from the Upper reach in 2017 was in the particulate fraction

***

```{r perc phg net upper bypass}
hg_net_upper_yb <- loads_net_clean17 %>% 
  filter(
    Reach == "Upper",
    AnalyteGroup == "Hg"
  ) %>% 
  group_by(Analyte) %>% 
  summarize(avg_net_load = mean(net_load))

perc_phg_net_upper_yb <- 
  hg_net_upper_yb$avg_net_load[2]/(hg_net_upper_yb$avg_net_load[1] + hg_net_upper_yb$avg_net_load[2])
```

Almost all of the net internal retention of Hg was due to the particulate fraction which accounted for `r round(perc_phg_net_upper_yb * 100)`% of the mercury lost in this reach 

***

```{r perc net liberty}
net_liberty <- loads_net_clean17 %>% 
  filter(
    Reach == "Liberty",
    AnalyteGroup %in% c("Hg", "MeHg")
  ) %>% 
  group_by(Analyte) %>% 
  summarize(avg_net_load = mean(net_load))

perc_fmehg_net_liberty <- 
  net_liberty$avg_net_load[1]/(net_liberty$avg_net_load[1] + net_liberty$avg_net_load[2])

perc_phg_net_liberty <- 
  net_liberty$avg_net_load[5]/(net_liberty$avg_net_load[4] + net_liberty$avg_net_load[5])
```

Overall, `r round(perc_fmehg_net_liberty * 100)`% of the MeHg produced in this reach in 2017 was in the filter-passing fraction, which is much different than what occurred in the upper reach and the entire Yolo Bypass 

Most of the average net source of mercury from the Liberty Island reach was in the particulate fraction at `r round(perc_phg_net_liberty * 100)`% 

